### 2011-06-16 paper/poster.tex
start poster and add missing file

\documentclass[landscape,a0paper,final,showframe]{baposter}

\usepackage{times}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{bm}

\usepackage{graphicx}
\usepackage{multicol}

\usepackage{pgfbaselayers}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

\newcommand{\captionfont}{\footnotesize}

\selectcolormodel{cmyk}

\graphicspath{{images/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Format 
\newcommand{\Matrix}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\Vector}[1]{\Matrix{#1}}
\newcommand*{\SET}[1]  {\ensuremath{\mathcal{#1}}}
\newcommand*{\MAT}[1]  {\ensuremath{\mathbf{#1}}}
\newcommand*{\VEC}[1]  {\ensuremath{\bm{#1}}}
\newcommand*{\CONST}[1]{\ensuremath{\mathit{#1}}}
\newcommand*{\norm}[1]{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand*{\abs}[1]{\mathopen| #1 \mathclose|}% use instead of $\|x\|$
\newcommand*{\absLR}[1]{\left| #1 \right|}% use instead of $\|x\|$

\def\norm#1{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand{\normLR}[1]{\left\| #1 \right\|}% use instead of $\|x\|$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{0.7em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\typeout{Poster Starts}
\definecolor{silver}{cmyk}{0,0,0,0.3}
\definecolor{yellow}{cmyk}{0,0,0.9,0.0}
\definecolor{reddishyellow}{cmyk}{0,0.22,1.0,0.0}
\definecolor{black}{cmyk}{0,0,0.0,1.0}
\definecolor{darkYellow}{cmyk}{0,0,1.0,0.5}
\definecolor{darkSilver}{cmyk}{0,0,0,0.1}

\definecolor{lightyellow}{cmyk}{0,0,0.3,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}
\begin{poster}{
  % Show grid to help with alignment
  grid=yes,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=lighteryellow,
  bgColorTwo=lightestyellow,
  borderColor=reddishyellow,
  headerColorOne=yellow,
  headerColorTwo=reddishyellow,
  headerFontColor=black,
  boxColorOne=lightyellow,
  boxColorTwo=lighteryellow,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=yes,
  headerborder=open,
  headerheight=0.08\textheight,
  headershape=roundedright,
  headershade=plain,
  headerfont=\Large\textsf, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  { % No eye catcher for this poster. If an eye catcher is present,
    % the title is centered between eye-catcher and logo.
    \includegraphics[height=5.5em]{figs/cavitysize-060}
  }
  % Title
  {\sf %Sans Serif
  %\bf% Serif
  A Classical Density Functional for Water}
  % Authors
  {\sf %Sans Serif
  % Serif
  David Roundy\hspace{3em}
  Jessica Hughes\hspace{3em}
  Oregon State University, USA
  }
  % University logo
  {
    \includegraphics[height=5.5em]{figs/osu-logo}
  }

  \tikzstyle{light shaded}=[top color=baposterBGtwo!30!white,bottom color=baposterBGone!30!white,shading=axis,shading angle=30]

  % Width of left inset image
     \newlength{\leftimgwidth}
     \setlength{\leftimgwidth}{0.78em+8.0em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}[1]{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=baposterBGone!80!black!#1!white,line width=0.03em] (0,0) circle(0.18em);}}

  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{ Classical
    Density-Functional Theory is a continuum empirical approach to
    modelling fluid behavior when the density is inhomogeneous---as is
    the case at interfaces.  We have developed a classical density
    functional for water, which provides a description of hydrophobic
    interactions.  }

  \headerbox{Statistical Associating Fluid
    Teory}{name=saft,column=0,below=cdft}{ The statistical associating
    fluid theory (SAFT) is a well-established approach for the
    description of the thermodynamics and phase equilibria of fluid
    systems, with particular applicability to hydrogen-bonding fluids.
    Our functional reduces in the homogeneous limit to the SAFT model
    for water developed in reference~\cite{clark2006developing}.
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
    \end{center}
    This functional accurately describes the liquid-vapor phase
    diagram away from the critical point, but gives a poor description
    of the compressibility.
  }

  \headerbox{The
    Functional}{name=fitting,column=0,below=saft,above=bottom}{ We
    introduce a new classical density functional for water.  The
    Helmholtz free energy functional is based on SAFT, and is composed
    of the usual terms:
    \begin{equation}
      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
    \end{equation}
    We model the repulsive interactions using the White Bear version
    of the Fundamental-Measure Theory~(FMT)
    functional~\cite{roth2002whitebear}.  The association term is
    taken from~\cite{yu2002fmt-dft-inhomogeneous-associating}, and we
    used a simple and efficient functional for the dispersion
    interaction with a single tunable parameter.
  }

  \headerbox{Surface Tension}{name=1D,column=1,row=0}{

  }
  \headerbox{A Cylinder in Water}{name=rod,column=1,below=1D,span=1,above=bottom}{
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/density-vs-radius}
    \end{center}
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
    \end{center}
  }
  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,row=0}{
    \begin{multicols}{2}
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
    \end{multicols}\vspace{-1em}
    \begin{center}
      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
      \hspace{1em}
      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
    \end{center}

    %      \begin{multicols}{2}
    {
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      %Note the different scales for the two graphs.
      %Our approach has a high accuracy on the neutral (UND) dataset.
    }
    %      \end{multicols}\vspace{-1em}
  }%
  \headerbox{Future Work}{name=future work,column=2,span=1,above=bottom,below=two rods}{

  }
  \headerbox{References}{name=references,column=3,above=bottom,below=two
  rods}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bibliographystyle{plain}	% (uses file "plain.bst")
    \bibliography{paper}
    %% \smaller
    %% \vspace{-0.4em}
    %% \bibliographystyle{ieee}
    %% \renewcommand{\section}[2]{\vskip 0.05em}
    %%   \begin{thebibliography}{1}\itemsep=-0.01em
    %%   \setlength{\baselineskip}{0.4em}
    %%   \bibitem{amberg07:nonrigid}
    %%     B.~Amberg, S.~Romdhani, T. Vetter.
    %%     \newblock {O}ptimal {S}tep {N}onrigid {ICP} {A}lgorithms for {S}urface {R}egistration
    %%     \newblock In {\em CVPR 2007}
    %%   \bibitem{amberg08:recognition}
    %%     B.~Amberg, R.~Knothe, T. Vetter.
    %%     \newblock Expression Invariant Face Recognition with a 3D Morphable Model
    %%     \newblock In {\em AFGR 2008}
    %%   \end{thebibliography}
  }%
\end{poster}%
%
\end{document}