### 2013-01-18 papers/polyhedra/paper.tex
add sketchy polyhedra paper Tested everything on darcy.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
\usepackage{graphicx}% Include figure files
%\usepackage{breqn}
\usepackage{color}

\newcommand{\fixme}[1]{\red{[#1]}}

\begin{document}
\title{Fundamental measure theory for hard polyhedra.}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}
%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We develop a classical density functional for hard polyhedra.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Why study hard polyhedra? Hard spheres are the standard reference
system for essentially all fluids, but always have 12-fold
coordination.  A few very interesting liquids have tetrahedral
coordination, which is particularly significant when we consider the
freezing transition.  By studying hard polyhedra, we potentially can
develop a reference functional that can incorporate orientational
effects.

There have been several studies of non-spherical hard shapes.  At
least one omitted the rotational degrees of freedom (parallel hard
cubes, add citation).  At least one included the rotational degrees of
freedom, treating each orientation as a distinct species, resulting in
a density that exists in six dimensions rather than three, which was
applied to systems with a high aspect ratio, which undergo a nematic
transition.  It seems a good thing to have a density functional that
is more efficient, with fewer degrees of freedom to optimize.  In
particular, we are optimistic that one can do better for ``almost
spherical'' polyhedra.

\section{Main ideas}

\newcommand\rr{\mathbf{r}}
\newcommand\rhat{\mathbf{\hat{r}}}
\newcommand\rot{\mathbf{\varpi}}

We will begin with the extension of fundamental measure theory (FMT)
to non-spherical particles by (so and so: add citation).  This theory
is based around a density that is a function of both position $\rr$
and orientation $\rot$.  The weighted densities are defined for a
given orientation, and then rotated appropriately when computing the
fundamental measures:
\begin{align}
  n_\alpha(\rr) &= \int n(\rr',\rot) w_\alpha(R(\rot)(\rr-\rr')) d\rr' d\rot
\end{align}
where $R(\rot)$ is the rotation operator corresponding to the
orientation $\rot$, and $w_\alpha(\rr)$ is the fundamental measure of
a single particle in the reference orientation at the origin.

Given these fundamental measures, the free energy is computed with
the standard Rosenfeld formula, plus a single correction, which we
will discuss later.
\begin{align}
  A_{HS} &= k_BT \int \Phi(\{n_\alpha(\rr)\}) d\rr
\end{align}

We propose to treat the orientational dependence of the fundamental
measures using an expansion in spherical harmonics:
\begin{align}
  w_\alpha(\rr) &= \sum_{lm} \tilde{w}_\alpha^{lm}(|\rr|) Y_{lm}(\rhat)
\end{align}
Given this expansion, we can simplify the weighted densities:
\begin{align}
  n_\alpha(\rr) &= \int \sum_{lm} n(\rr',\rot)
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm}(\widehat{R(\rot)(\rr-\rr')}) d\rr' d\rot
  \\
  &= \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm'}(\widehat{(\rr-\rr')})
  n(\rr',\rot)
       D_{lmm'}(\rot) d\rot d\rr'
       \\
  &=
    \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
    Y_{lm'}(\widehat{(\rr-\rr')})
    n_{lmm'}(\rr') d\rr'
\end{align}
This reduces the challenge of handling the rotation to that of solving
for a set of densities $n_{lmm'}(\rr)$.
\subsection{Ideal gas free energy}
To address this, we will
first consider the free energy of an ideal gas.
\begin{align}
  A_{\textit{ideal}} &=
  k_BT \int
  \left(n(\rr,\rot)\ln(\Lambda^3\Theta^3 n(\rr,\rot)) - n(\rr,\rot)
  \right) d\rr d\rot
  \\
  &=
  k_BT \sum_{lmm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)\left(
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right) - 1
  \right) d\rr d\rot
  \\
  &=
  k_BT \int n_{000}(\rr)
  \left(\ln\left(\Lambda^3\Theta^3 \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)\right) - 1
  \right) d\rr d\rot
  \\&+
  k_BT \sum_{l>1,mm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &= A_{000} + \sum_{l>1,mm'} A_{lmm'} - k_BT\int n_{000}(\rr)d\rr
\end{align}
Here I have broken the $000$ term out of the outer summation over
$lmm'$, since this one is special, since it is independent of $\rot$.
We'll tackle that integral first, since it is a tad simpler.
\begin{align}
  A_{000} &= k_BT\int
  n_{000}(\rr)\ln\left(
  \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)
  \right)
  d\rr d\rot
\end{align}
We can simplify this integral a bit if we write the log as a power
series...
\begin{align}
  A_{000}
  &= k_BT\int
  n_{000}(\rr)\ln\left(
  n_{000}(\rr)\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  \right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\ln\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\&-\frac{k_BT}{2!}\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)^2
  d\rr d\rot
  \\&+\cdots
\end{align}
Now the linear term in the power series will vanish, since the
summation is over $l>0$ only.  In the quadratic term, only a few terms
will survive, and we'll end up with
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} n_{lmm'}(\rr)n_{l,-m,-m'}}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} |n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
\end{align}
Now an interesting question is whether we can mimic this power series
with a log, such as
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln \left( \sqrt{n_{000}(\rr)^2 + \sum_{lmm'}
    c_ln_{lmm'}(\rr)^2} \right)d\rr
  \\
  &= 
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int n_{000}(\rr)\ln\left( \sqrt{1 + \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)^2}} \right)d\rr
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int \frac12 \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)} d\rr
  \\&\quad+\cdots
\end{align}
from which we can conclude that to get an answer correct to the
quadratic order, we need
\begin{align}
  c_l &= -\frac{2l+1}{8\pi^2}
\end{align}
which seems all right except for the sign:  I expect that the free
energy should increase as we increase $n_{lmm'}$, since it means the
system is becoming more ordered, and it looks like to this quadratic
term, the free energy decreases... except that we're only looking at
$A_{000}$ right now, and thus are neglecting a possibly significant
increase due to the other terms...  so perhaps we should call this all
right, and move on for now.

The $A_{lmm'}$ terms are a bit different, but not too different:
\begin{align}
  A_{lmm'} &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \ln\left(\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \ln\left(1 + \frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \left(\frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}+\cdots\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) \frac{8\pi^2}{2l+1} \frac{n_{l,-m,-m'}(\rr)}{n_{000}(\rr)}
  d\rr + \cdots
  \\
  &=
  k_BT \int \frac{8\pi^2}{2l+1} \frac{|n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr + \cdots
\end{align}
It is interesting how similar the $A_{lmm'}$ terms are to the
quadratic terms in $A_{000}$, off by a factor of $-\frac12$.  This
leads me to wonder about the possibility of writing
\begin{align}
  A_{\textit{ideal}} &\approx
  k_BT \int n_{000}(\rr)\left( \ln\left(
  \sqrt{\sum_{lmm'}\frac{2l+1}{16\pi^2}|n_{lmm'}|^2}
  \right) - 1\right)d\rr
\end{align}
This would be correct up to the quadratic terms, and doesn't require
any rotational integrals.  It also seems liable to be reasonable up to
higher order terms, although I have no guarantee of that.  An
interesting excersize would be to take this out to the fourth order,
and also take the exact expression out to the fourth order and see how
they compare.  They will actually differ in the third order terms, I
expect, since the exact expression will indeed have nonzero third
order terms.

So we have two plausible options that I can see: we could either go
with this logarithmic approximation, or we could just go out to second
order.  The latter is probably more defensible in terms of
perturbation theory, and would give us something like
\begin{align}
  A_{\textit{ideal}} &\approx
  k_BT \int \left(n(\rr)\ln n(\rr) - n(\rr) +
  \sum_{lmm'}\frac{2l+1}{16\pi^2}\frac{|n_{lmm'}|^2}{n_{000}(\rr)}
  \right)d\rr
\end{align}
where I have substituted the notation $n(\rr)$ for $n_{000}(\rr)$ to
save space.

\subsection{Finding the orientational moments}
We find the orientational moments $n_{lmm'}(\rr)$ by minimizing the
free energy.  Let's begin by optimizing the free energy with respect
to nonzero angular momenta:
\begin{align}
  0&=\frac{\delta A}{\delta n_{lmm'}(\rr)} \\
  &= \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  + k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \frac{\delta n_\alpha(\rr')}{\delta n_{lmm'}(\rr)} d\rr'
  \\
  &=
   \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  \\&\quad+ k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
Solving for $n_{lmm'}(\rr)$, we find that
\begin{align}\label{eq:nlmm'}
  n_{lmm'}(\rr) &= \frac{8\pi^2k_BT}{2l+1}
  \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
I note that this equation for $n_{lmm'}$ is approximate, since we made
an approximation when we neglected higher order terms in the expansion
of the log in the ideal gas free energy.  Thus in cases where there is
relatively little orientational ordering, it ought to be an accurate
equation.  In any case I imagine it will be plausible.

\subsection{The Procedure}
Our input when computing the free energy is the ordinary density
$n(\rr)$, which is precisely equal to $n_{000}(\rr)$ above, because it
is just the orientationally averaged density, which is also what
$n_{000}(\rr)$ is.  From this density, we can compute a first set of
fundamental measures
\begin{align}
  n_\alpha(\rr) &= \int \tilde{w}_\alpha^{000}(|\rr-\rr'|)n(\rr')d\rr'
\end{align}
where we will discuss elsewhere how we find the weighting functions
$\tilde{w}_\alpha^{000}$.  Thus we have our zeroth-order approximation
to the fundamental measures.  At this point we can compute a
zeroth-order free energy just by evaluating the free energy functional
with these fundamental measures, which would be the true free energy
if there were no orientational ordering.

From the zeroth-order fundamental measures, we can
compute the orientational densities using Eq.~\ref{eq:nlmm'}.  Once we
have these densities, we can find an improved set of fundamental
measures:
\begin{align}
  n_\alpha(\rr) &= \sum_{lmm'} \int
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|) Y_{lm'}(\rr-\rr')n_{lmm'}(\rr')
\end{align}
With this improved set of fundamental measures, we can now just
evaluate the free energy.  How could it be simpler? Well, it could be
more complicated if we were to go back and try to compute an improved
set of $n_{lmm'}(\rr)$ using our improved fundamental measures.  I
suspect this would be silly, since we have already dropped out cubic
terms in $A_{\textit{ideal}}$.  No benefit from precisely computing a
higher order correction to a low-order approximation.  It \emph{is}
worth checking (particularly while debugging) that our two
approximations for the fundamental measures aren't \emph{too} wildly
different, and that the two approximations to the free energy also
aren't too wildly different.

\subsection{Finding the weighted densities}

\newcommand\xhat{\mathbf{\hat{x}}}
\newcommand\yhat{\mathbf{\hat{y}}}
\newcommand\zhat{\mathbf{\hat{z}}}
\newcommand\facehat{\mathbf{\hat{f}}}
\newcommand\Rface{\ensuremath{R_{\text{face}}}}
\newcommand\Rvertex{\ensuremath{R_{\text{vertex}}}}

Here I will discuss how to find the weighted densities we need in
order to study polyhedra.  We begin with $n_3(\rr)$, the packing
fraction.
\begin{align}
  w_3(\rr) &= 1 \text{ inside the polyhedron}\\
  &= 0 \text{ outside the polyhedron} \\
  &= \prod_i^{\text{faces}} \Theta(\Rface - \rr\cdot\facehat_i)
\end{align}
This is pretty easy, although the integrals will be less so.  The
$000$ integral is always going to give us something that is
essentially the same, but I won't go into this yet.

$n_2(\rr)$ is the area, which we can easily get from the above
weighted density.
\begin{align}
  w_2(\rr) &= |\nabla w_3(\rr)| \\
  &= \sum_i^{\text{faces}} \delta(\Rface-\rr\cdot\facehat_i)
  \prod_{j\ne i} \Theta(\Rface-\rr\cdot\facehat_j)
\end{align}
We could also just ask what the area is at a given position.

$n_1(\rr)$ is the first fundamental measure that is interestingly
different for polyhedra than it is for spheres.  This is a measure of
the mean curvature, and polyhedra have no curvature except at the
edges.  Thus $n_1(\rr)$ is going to be nonzero only on the edges, and
there it will be a delta function.
\begin{align}\label{eq:w1}
  w_1(\rr) &= \sum_{ij}^{\text{NN faces}}
  \Theta\left(\Rvertex - r\right)
  \delta(\Rface - \rr\cdot \facehat_i)
  \delta(\Rface - \rr\cdot \facehat_j)
\end{align}
where $\facehat_i$ is a unit vector in the direction of the $i$ face
of the polyhedron, $\Rvertex$ is the distance between the center of
the polyhedron and the vertex, and $\Rface$ is the distance of the
center of the face from the origin.  WARNING: I may have something off
in these formulas, but the general idea is there.  To find a constant
factor, we'd consider the

$n_0(\rr)$ is the fundamental measure of gaussian curvature, which is
the product of the curvature in the two eigendirections.  This is zero
(for polyhedra) except at the corners, so $n_0(\rr)$ itself is going
to be zero except at the corners.  Also, the gaussian curvature
must satisfy a sum rule, which tells us that
\begin{align}
  w_0(\rr) &= \frac1{N}\sum_{i}^{N\text{ vertices}} \delta(\rr-\rr_i)
\end{align}
Interestingly, this meas that
\begin{align}
  w_0^{000}(\rr) &= \frac{\delta(\Rvertex - |\rr|)}{4\pi \Rvertex^2}
\end{align}
which means that as far as $w_0^{000}$ is concerned, the polyhedron is a
sphere with a radius equal to the vertex distance (i.e. the distance
from the center to the vertex of the polyhedron).

\subsection{Tetrahedron}

\subsubsection{$w_3^{000}$}

In order to solve for $w_3^{000}(r)$ of a tetrahedron, we need to
consider where $r$ lies.  If $r<\frac{\sqrt{3}}{2}R$ (which is
\Rface), then the sphere lies entirely within the tetrahedron, and
$w_3^{000}(r)=1$.  Conversely, if $r<\sqrt{3}R$ (which is \Rvertex),
then $w_3^{000}(r)=0$.  There are two in-between cases:  either
$R<r<\Rvertex$ (in which case each vertex forms an isolated triangle
poking out of the sphere), or $\Rface<r<R$, in which case isolated
circles of spherical surface poke out of the faces of the
tetrahedron.  The latter case is pretty easy, if we just integrate
over the sticking-out bits, and subtract that from the total surface
area of the sphere:
\begin{align}
  w_3^{000}(r) &= 1 - \frac1{\pi}\int d\Omega \\
  &= 1 - \frac1{\pi}2\pi \int_0^{\cos^{-1}\left(\frac{\sqrt{3}R}{2r}\right)} \sin\theta d\theta
  \\
  &= 1 - 2 \int^1_{\frac{\sqrt{3}R}{2r}} du
  \\
  &= \frac{\sqrt{3}R}{r} - 1
  \\
  \frac{\sqrt{3}R}{2}&<r<R
\end{align}

The final consideration (for $w_3^{000}(r)$ of a tetrahedron) is when
$R < r < \sqrt{3}R$, when we will want to just integrate over the
curvy triangles formed by the portion of the tetrahedron that stick
through the sphere.  We just need to integrate through one of these
points, and multiply by four... which might be easiest to do in a
Cartesian coordinates rotated by 45$^\circ$ around the ordinary $z$
axis.
\begin{align}
  w_3^{000}(r)
  &=
  \frac4{4\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_0^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  u&= \sqrt{x^2+y^2+z^2}\\
  du &= \frac{ydy}{\sqrt{x^2+y^2+z^2}}
  \\
  w_3^{000}(r)
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  &=
  \frac1{\pi r^2}
  \int_{\sqrt{r^2-2R^2}}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  z^2+x^2+y^2 &= r^2 \\
  z_{min}^2 + (R-z_{min})^2/2 &= r^2
\end{align}
Instead I'm going now to think of using cylindrical coordinates...
\begin{align}
  w_3^{000}(r) &= 3\frac{4}{4\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  u &= z^2+s^2 \\
  du &= sds \\
  &= \frac{\sqrt{u^2-z^2}}{u}ds
  \\
  w_3^{000}(r)
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_{z}^{\sqrt{z^2+\left(2\frac{\sqrt{3}R-z}{\cos\phi}\right)^2}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2}}
  \\
  R &= \frac{2}{\sqrt{3}}s\cos\phi + \frac{1}{\sqrt{3}}z \\
  s &= 2\frac{\sqrt{3}R-z}{\cos\phi}
\end{align}
\newcommand\thetamax{\theta_{\text{max}}}
\begin{widetext}
Maybe spherical coordinates? Rotated $\pi/4$? Let's look first at the
plane that is our limit for $\theta$:
\begin{align}
  \sqrt{3}R &= x + y + z \\
  &= \sqrt{2}r\sin\thetamax\cos\phi +
  r\cos\thetamax
  \\
  \sqrt{3}\frac{R}{r} &=
  \sqrt{2}\sqrt{1-\cos^2\thetamax}\cos\phi
  + \cos\thetamax
  \\
  \sqrt{1-\cos^2\thetamax} &=
  \frac{\sqrt{3}\frac{R}{r}-\cos\thetamax}{\sqrt{2}\cos\phi}
  \\
  1-\cos^2\thetamax &=
  \frac{3\frac{R^2}{r^2}+\cos^2\thetamax -
    2\sqrt{3}\frac{R}{r}\cos\thetamax}{2\cos^2\phi}
  \\
  0 &=
  (1+2\cos^2\phi)\cos^2\thetamax
  - 2\sqrt{3}\frac{R}{r}\cos\thetamax
  + \left(3\frac{R^2}{r^2} - 2\cos^2\phi\right)
  \\
  \cos\thetamax &= \frac{2\sqrt{3}\frac{R}{r}
  - \sqrt{12\frac{R^2}{r^2} - 4(1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{2(1+2\cos^2\phi)}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (2-\cos2\phi)\left(3\frac{R^2}{r^2}-1+\cos2\phi\right)}}{2-\cos2\phi}
\end{align}
\end{widetext}
This is sort of a mess, but does look useable.
\begin{align}
  w_3^{000}(r) &=
  \frac{12}{4\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi
  \int_0^{\thetamax}d\theta
  \sin\theta
  \\
  &=
  \frac{3}{\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}
  (1 - \cos\thetamax)d\phi
\end{align}
Surely we can either use Euler's relation or a geometrical
construction to solve for $\thetamax$?!

\subsubsection{$w_2^{000}$}
The easiest way to get $w_2^{000}(r)$ is just to take a derivative of
$w_3^{000}(r)$ we should have just computed:
\begin{align}
  w_2^{000}(r) &= \frac{dw_3^{000}(r)}{dr}
\end{align}

\subsubsection{$w_1^{000}$}
As mentioned before, $w_1^{000}(r)$ is the average of an integral
around the edges of the tetrahedron.  This is a little bit tricky, as
seen in Eq.~\ref{eq:w1}.  $w_1^{000}$ is zero except when $\sqrt{2}R <
r < \sqrt{3}R$.
\begin{align}
  \rr &= r\left(
  \sin\theta\cos\phi\xhat +
  \sin\theta\sin\phi\yhat +
  \cos\theta\zhat \right)
  \\
  \facehat_1 &= \frac{\xhat+\yhat+\zhat}{\sqrt{3}} \\
  \facehat_2 &= \frac{-\xhat-\yhat+\zhat}{\sqrt{3}} \\
  \rr\cdot\facehat_1 &=
  \frac{r}{\sqrt{3}}\left(
  \cos\theta+\sin\theta\cos\phi + \sin\theta\sin\phi
  \right)
  \\
  &= \frac{r}{\sqrt{3}}\left(
  \cos\theta+\sqrt{2}\sin\theta\cos(\phi+\pi/4)
  \right)
  \\
  \rr\cdot\facehat_2 &=
  \frac{r}{\sqrt{3}}\left(
  \cos\theta-\sqrt{2}\sin\theta\cos(\phi+\pi/4)
  \right)
\end{align}
\begin{widetext}
\begin{align}
  w_1^{000}(r) &= \frac{6}{4\pi}\int
  \delta(\sqrt{2}R - \rr\cdot \facehat_1)
  \delta(\sqrt{2}R - \rr\cdot \facehat_2)
  \sin\theta d\theta d\phi
  \\
  &= \frac{6}{4\pi}\int
  \delta\left(\sqrt{2}R
  - \frac{r}{\sqrt{3}}(\cos\theta-\sin\theta\cos(\phi+\pi/4))\right)
  \delta\left(\sqrt{2}R
  - \frac{r}{\sqrt{3}}(\cos\theta+\sin\theta\cos(\phi+\pi/4))\right)
  \sin\theta d\theta d\phi
\end{align}
\end{widetext}
\begin{align}
  u &= \sqrt{2}R -\frac{r}{\sqrt{3}}(\cos\theta + \sin\theta\cos(\phi+\pi/4)) \\
  du &= \frac{r}{\sqrt{3}}\sin\theta\sin(\phi+\pi/4)d\phi \\
  d\phi &= \frac{\sqrt{3}}{r\sin\theta}\frac{1}{\sin(\phi+\pi/4)}
  \\
  w_1^{000}(r)
  &= \frac{6}{4\pi}\int
  \delta\left(-2\frac{r}{\sqrt{3}}\cos\theta+u\right)
  \delta(u)
   \frac{\sqrt{3}}{r}\frac{ d\theta du}{\sqrt{1-u^2}OOPS!}
  \\
  &= \frac{6}{4\pi}\int
  \delta\left(-2\frac{r}{\sqrt{3}}\cos\theta\right)
   \frac{\sqrt{3}}{r}d\theta
   \\
  v &= -2\frac{r}{\sqrt{3}}\cos\theta \\
  dv &= 2\frac{r}{\sqrt{3}}\sin\theta d\theta \\
  &= 2\frac{r}{\sqrt{3}}\sqrt{1-v^2}d\theta \\
  w_1^{000}(r)
  &= \frac{6}{4\pi}\int
  \delta\left(v\right)
   \frac{6}{r^2}\frac{dv}{\sqrt{1-v^2}} \\
   &= \frac{9}{\pi r^2}
\end{align}
I'm a little doubtful of the above angular integral.  I don't see how
it will be able to integrate to anything very reasonable when
integrating over $r$.  In particular, I am pretty certain that
\emph{has} to depend on the distance of closest approach.
I know that
\begin{align}
  \int_{r_{min}}^{r} w_1^{000}(r)4\pi r^2dr
  &\propto \sqrt{r^2 - r_{min}^2}
\end{align}

\subsection{Cube}

In calculating $w_3^{000}$ we only need to integrate over one octant.
\begin{align}
  w_3^{000}(r) &= \int \Theta(R-x)\Theta(R-y)\Theta(R-z) d\Omega
  \\
  &= 8\int_0^{\frac{\pi}{2}} \int_0^{\frac{\pi}{4}}
  \Theta(R-r\sin\theta\cos\phi)
  \Theta(R-r\sin\theta\sin\phi)
  \Theta(R-r\cos\theta)
  r^2\sin\theta d\theta d\phi
  \\
  &=
  x
\end{align}
There are a couple of possibilities.  If $r<R$ then $w_3^{000}(r) =
1$.  If $r>\sqrt{3}R$ the $w_3^{000}(r) = 0$.  That leaves two
interesting possibilities.  If $R<r<\sqrt{2}R$ then there are six
edges to the integral, and we find that
\begin{align}
  w_3^{000}(r) &= \frac1{4\pi r^2}\int \Theta(R-x)\Theta(R-y)\Theta(R-z) d\Omega
\end{align}
If $\sqrt{2}R<r<\sqrt{3}R$ then there are just three sides to the
integral, as illustrated in Fig.~\ref{fig:rbig}.  This integral is
easiest to do by integrating over just one of the eight protruding
corners.  This seems easiest if we define the corner as the $z$ axis.
\begin{align}
  w_3^{000}(r) &= \frac{6}{4\pi}
  \int_{0}^{\frac{\pi}{3}}
  \int_0^{\theta_{m}} \sin\theta d\theta d\phi \\
  r\cos\theta_{m} + r\sin\theta_{m}\cos\phi  &= R
  \\
  \cos\theta_{m} + \sin\theta_{m}\cos\phi  &= \frac{R}{r} \\
  w_3^{000}(r) &= \frac{6}{4\pi}
  \int_{0}^{\frac{\pi}{3}}
  \int_0^{\theta_{m}} \sin\theta d\theta d\phi
\end{align}
\begin{align}
  w_3^{000}(r) &= \frac{8}{4\pi r^2}\iint_0^R \delta(r-\sqrt{x^2+y^2+z^2})
  dxdydz \\
  u &= \sqrt{x^2+y^2+z^2} \\
  du &= \frac{zdz}{u} \\
  w_3^{000}(r) &= \frac{8}{4\pi r^2}\iint_0^R \delta(r-u)
  dxdy\frac{udu}{\sqrt{u^2 -x^2-y^2}} \\
  &= \frac{2}{\pi r}\int_{r_0}^R\int_{r_0}^R
  dxdy\frac{1}{\sqrt{r^2 -x^2-y^2}} \\
  &= \frac{2}{\pi r}\int_{\sqrt{r^2-2R^2}}^R\int_{\sqrt{r^2-R^2-x^2}}^R
  \frac{dxdy}{\sqrt{r^2 -x^2-y^2}} \\
\end{align}
\begin{figure}
\includegraphics[width=2in]{figs/w3cube}
\caption{Picture of the sphere with $\sqrt{2}R<r<\sqrt{3}R$.}\label{fig:rbig}
\end{figure}

%\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}