### 2013-02-07 papers/polyhedra/harmonics.tex
add empty paper about harmonics Tested everything on node08.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}

\usepackage{color}

\begin{document}
\title{Harmonics with symmetries of platonic solids}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}
%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This is just some notes to myself regarding harmonics.
\end{abstract}

\newcommand\rot{\ensuremath{\mathbf{\varpi}}}
\newcommand\rhat{\ensuremath{\mathbf{\hat{r}}}}
\newcommand\solidangle{\ensuremath{\Omega}}

\maketitle

I plan on adding here an introduction to symmetrized linear
combinations of spherical harmonics with the symmetries of platonic
solids.

\end{document}