### 2013-08-23 papers/water-saft/paper.tex

\documentclass[twocolumn,amsmath,amssymb,prl]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Definitions
\def\re{\text{Re}}
\def\im{\text{Im}}
\def\ket#1{\vert #1 \rangle}
\def\cU{{\cal{U}}}
\def\cD{{\cal{D}}}
\def\re{\text{Re}}
\def\im{\text{Im}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\xx}{\textbf{r}}
\newcommand{\refnote}{\red{[ref]}}

\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
\newcommand\kappaassoc{\ensuremath{\kappa_\textit{a}}}
\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
\newcommand\lscale{\ensuremath{s_d}}

% fixme is intended to prioritize stuff that needs fixing.
\newcommand{\fixme}[1]{\textcolor{red}{[\emph{#1}]}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{A Classical Density-Functional Theory for Describing Water Interfaces}

\author{Eric J. Krebs}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{Jeff B. Schulte}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}

\begin{abstract}
We develop an improved version of our old functional.
\end{abstract}
\maketitle

\section{Introduction}

Water, the universal solvent, is of critical practical importance, and
a continuum description of water is in high demand for a solvation
model.  A number of recent attempts to develop improved solvation
models for water have built on the approach of classical
density-functional theory (DFT)~\cite{jeanmairet2013molecular,
  zhao2011molecular, zhao2011new, ramirez2005direct,
  ramirez2005density, levesque2012solvation, levesque2012scalar}.
Classical DFT is based on a description of a fluid written as a free
energy functional of the density distribution.  There are two general
approaches used to construct a classical DFT for water.  The first is
to choose a convenient functional form which is then fit to properties
of the bulk liquid at a given temperature and pressure
\cite{jeanmairet2013molecular, lischner2010classical} ~\fixme{cite
  Jeanmairet and similar}.  Using this approach, it is possible to
construct a functional that reproduces the exact second-order response
function of the liquid under the fitted conditions.  However, this
class of functional will be less accurate at other temperatures or
pressures---and in the inhomogeneous scenarios in which solvation
models are applied.  The second approach is to construct a functional by
applying liquid-state theory to a model system, and then fit the model
to experimental data such as the equation of
state~\cite{hughes2013classical, clark2006developing, gloor2002saft,
  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination,
  sundararaman2012computationally}.  Classical DFT is a natural
framework for creating a more flexible theory of hydrophobicity that
can readily describe interaction of water with arbitrary external
potentials---such as potentials describing strong interactions with
solutes or surfaces.

A widely used family of models used in the development of classical
density functionals is based on Statistical Associating Fluid Theory
(SAFT)~\cite{chapman1989saft}.  SAFT is a theory based on a model of
hard spheres with weak dispersion interactions and hydrogen-bonding
association sites, which has been used to accurately model the
equations of state of both pure fluids and mixtures over a wide range
of temperatures and pressures~\cite{muller2001molecular,
  tan2008recent}.  The association term in SAFT is based on Wertheim's
first-order thermodynamic perturbation theory, which models fluids as
hard-spheres with strong associative interactions at specific sites on
the surfaces of the spheres~\cite{wertheim1984fluidsI,
  wertheim1984fluidsII, wertheim1986fluidsIII, wertheim1986fluidsIV}.
These association sites have an attractive interaction at contact, and
rely on the hard-sphere radial distribution function at contact
$g_\sigma^\text{HS}$ in order to determine the extent of association.
While this function is well-known for the homogeneous hard-sphere
fluid, it is less well-known for inhomogeneous systems, such as occur
at liquid interfaces.  In a recent paper, we examined this function at
a variety of hard-sphere interfaces~\cite{schulte2012using}.  We
tested the accuracy of existing approximations for the distribution
function at contact~\cite{yu2002fmt-dft-inhomogeneous-associating,
  gross2009density}, and derived a significantly improved
approximation for the averaged distribution function at contact.

We will apply this improved $g_\sigma^{HS}$ to a SAFT-based classical
density functional for water~\cite{hughes2013classical}.  This
functional was constructed to reduce in the homogeneous limit
to the 4-site optimal SAFT model for water developed by Clark \emph{et
  al.}~\cite{clark2006developing}.  The DFT of
Ref.~\citenum{hughes2013classical} uses the association free energy
functional of Yu and
Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is based on a
$g_\sigma^{HS}$ that we have since found to be
inaccurate~\cite{schulte2012using}.  In this paper, we will examine
the result of using the improved functional for $g_\sigma^{HS}$
developed in Ref.~\citenum{schulte2012using} to construct an
association free energy functional.

\section{Method}

The classical density functional for water from
Reference~\citenum{hughes2013classical} consists of four terms:
\begin{align}
  F[n(\rr)] &= F_{\text{ideal}}[n(\rr)] + F_{\text{HS}}[n(\rr)]
  + F_{\text{disp}}[n(\rr)] + F_{\text{assoc}}[n(\rr)]
\end{align}
where $F_{\text{ideal}}$ is the ideal gas free energy and
$F_{\text{HS}}$ is the hard-sphere excess free energy, for which we
use the White Bear functional~\cite{roth2002whitebear}.
$F_{\text{disp}}$ is the free energy contribution due to the
square-well dispersion interaction.  The dispersion term of the free
energy contains one empirical parameter, which is used to fit the
surface tension of water near one atmosphere.  Finally,
$F_{\text{assoc}}$ is the free energy contribution due to association,
which is the term that we examine in this paper.

The association free energy for our four-site model has the form:
\begin{align}
  F_\text{assoc}[n] &= k_BT \int n_\text{site}(\xx)
  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
\end{align}
where $n_\text{site}(\rr)$ is the density of bonding sites at
position~$\rr$:
\begin{align}
  n_\text{site}(\rr) &=
  \begin{cases}
    4 n(\rr) & \text{this work}\\
    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
  \end{cases}
\end{align}
where the factor of four comes from the four hydrogen bond sites,
$n_0(\rr)$ is the average density contacting point $\rr$, and
$\zeta(\xx)$ is a dimensionless measure of the density inhomogeneity
from Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  The
functional $X(\rr)$ is the fraction of association sites \emph{not}
hydrogen-bonded, which is determined for our 4-site model by the
quadratic equation
\begin{align}
  X(\xx) &= \frac{\sqrt{1 + 2n_\text{site}'(\rr)
      \kappaassoc g^\textit{SW}_\sigma(\xx)
  \left(e^{-\beta\epsilonassoc} - 1\right)} - 1}
  {n_\text{site}'(\rr)
    \kappaassoc g^\textit{SW}_\sigma(\xx)
  \left(e^{-\beta\epsilonassoc} - 1\right)}, \label{eq:X}
\end{align}
where
\begin{align}
  n_\text{site}'(\rr) &=
  \begin{cases}
    \frac{4}{\pi\sigma^2} \int n(\rr')\delta(\sigma - |\rr-\rr'|) d\rr' & \text{this work}\\
    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
  \end{cases} 
\end{align}
is the density of bonding sites that could bond to the sites~$n_\text{site}(\rr)$, and
\begin{align}
  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
\end{align}
%This is equation 77 of gil-villegas-1997-SAFT-VR
where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
contact for a hard-sphere fluid with a square-well dispersion
potential, and $a_1$ and $a_2$ are the two terms in the dispersion
free energy.  The radial distribution function of the square-well
fluid $g^\textit{SW}_\sigma$ is written as a perturbative correction
to the hard-sphere fluid correlation function $g^\textit{HS}_\sigma$.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/surface-tension}
\end{center}
\caption{Comparison of Surface tension versus temperature for
  theoretical and experimental data. The experimental data is taken
  from NIST~\cite{nistwater}.  The length-scaling parameter $\lscale$
  is fit so that the theoretical surface tension will match the
  experimental surface tension near room temperature.}
\label{fig:surface-tension}
\end{figure}

As in Ref.~\citenum{hughes2013classical}, we use Clark's five
empirical parameters, and fit the calculated surface tension to
experimental surface tension at ambient conditions by tuning the
parameter $\lscale$, which adjusts the length-scale of the average
density used for the dispersion interaction.  With the improved
association term, we find these agree when $\lscale$ is 0.454, which
is an increase from the value of 0.353 found in
Ref.~\citenum{hughes2013classical}.  In order to explore further the
change made by the improved association term, we compared the new
functional with that of Ref.~\citenum{hughes2013classical} for the two
hydrophobic cases of hard rods and the hard spherical solutes.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-compare}
\end{center}
\caption{ Density profiles for a water around a single hydrophobic rod
  of radius 0.1 nm. The solid red profile is from the functional
  developed in this paper and the dashed blue profile is the result
  from Ref. \cite{hughes2013classical}.  For scale, under the profiles
  is a cartoon of a string of hard spheres touching in one
  dimension. The horizontal black dotted line is the bulk density for
  water and the vertical line on the left represents the wall of the rod.}
\label{fig:density-single-rod}
\end{figure}

\section{Results}

We will first discuss the case of a single hydrophobic rod immersed in
water. Figure~\ref{fig:density-single-rod} shows the density profile
of water near a rod with radius~1~\AA.  The density computed using the functional of this
paper is qualitatively similar to that from
Ref.~\citenum{hughes2013classical}, with a comparable density at
contact---consistent with having made only a moderate change in the
free energy.  The first density peak near the surface is consistently
higher than that from Ref.~\citenum{hughes2013classical}, and the peak
is not smooth.  This results from the improved accuracy of the
$g_\sigma^\textit{HS}$ from Ref.~\citenum{schulte2012using}, since beyond the
first peak water molecules are unable to touch---or hydrogen bond
to---molecules at the surface of the hard rod.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/single-rod-broken-HB}
\end{center}
\caption{Broken hydrogen bonds per nanometer for hydrophobic rods
  immeresed in water.  The solid red line uses the functional developd
  in this paper while the dashed blue line uses the functional from
  Ref. \cite{hughes2013classical}. For large enough rods, the graph
  increases linearly for both functionals.}
\label{fig:single-rod-broken-HB}
\end{figure}

A common test case for studying hydrophobic solutes in water is the
hard-sphere solute.  Figure~\ref{fig:spheres-broken-HB} shows results
for the number of broken hydrogen bonds caused by a hard-sphere soute,
as a function of the solute radius.  As in
Fig.~\ref{fig:single-rod-broken-HB}, the number of broken bonds scales
with surface area for large solutes, and the number of broken bonds is
about four times smaller than the number from the functional of
Ref.~\citenum{hughes2013classical}.  For solutes smaller than
3~\AA\ in radius, there is less than a tenth of a hydrogen bond
broken. This is consistent with the well-known fact that small solutes
(unlike large solutes) do not disrupt the hydrogen-bonding network of
water~\cite{chandler2005}.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/sphere-broken-HB}
\end{center}
\caption{Broken hydrogen bonds for hard spheres immeresed in water.
  The solid red line uses our the functional developd in this paper
  while the dashed blue line is from
  Ref. \cite{hughes2013classical}l.}
\label{fig:spheres-broken-HB}
\end{figure}

\section{Conclusion}

We have modified the classical DFT for water developed by
Hughes~\emph{et al.}~\cite{hughes2013classical} with the more accurate
radial distribution function at contact developed by Schulte~\emph{et
  al.}~\cite{schulte2012using}, which affects the predicted hydrogen
bonding between water molecules.  We found that while this
modification has a relatively mild effect on the free energy and
density profiles, it predicts fewer broken hydrogen bonds around
hydrophobic solutes and at aqueous interfaces.

\bibliography{paper}% Produces the bibliography via BibTeX.

\clearpage

\section{Figures that are not included in the paper (but could be)}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/single-rod-X-plot}
\end{center}
\caption{ The average number of hydrogen bonds per molecule at
  distances from the center of hydrophobic rods of various
  radii. The solid lines are the results from using the assymetric
  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}^{HS}$.}
\label{fig:single-rod-X}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/spheres-X-plot}
\end{center}
\caption{ The average number of hydrogen bonds per molecule at
  distances from the center of spherical cavities of various
  radii. The solid lines are the results from using the assymetric
  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}$.}
\label{fig:spheres-X}
\end{figure}

\end{document}