### 2013-09-01 papers/pair-correlation/presentation/presentation.tex
Added a presentation folder in papers/pair-correlation. It contains a python program that generates a pdfs of a sample monte carlo simulation, and a latex file that puts them together. Run presentation/mc-circle-slow.py to get the frames and then compile presentation/presentation.tex for the finished animation. Tested everything on paho-deb.

\documentclass{beamer}
%\usepackage{beamerthemelined}
\usepackage{pstricks}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{graphicx}
\usepackage[draft]{animate}
%\setbeamertemplate{navigation symbols}{}
\beamertemplatenavigationsymbolsempty

\usetheme{Boadilla}
\usecolortheme{whale}
\setbeamertemplate{itemize items}[triangle]

\title{}
\author{}
\date{}

\begin{document}

\begin{frame}
  \frametitle{Monte Carlo Example}
  \begin{figure}[h]
    \centering
    \animategraphics[width=100mm,autoplay,final]{.6}{anim/mc-slow-}{000}{100}
  \end{figure}
\end{frame}

\end{document}