### 2014-03-22 papers/square-well-liquid/paper.tex
automagically generate two stupid plots for square-well paper (to demonstrate how) Tested on darcy. Took 0:20.07.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{color}

\begin{document}
\title{Classical density functional theory for the square-well liquid}

\author{Student One}
\author{Student Two}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

\begin{abstract}
  We have developed a functional.
\end{abstract}

\maketitle

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}
  \caption{A simple DOS plot!}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
  \caption{A simple DOS plot!}
\end{figure}

\end{document}