### 2014-05-21 papers/thesis-roth/project.tex
comments on Roth project report Tested on bennet. Took 0:33.33.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Definitions

% 1/k_B/T
\newcommand{\kT}{\ensuremath{k_BT}}

% vector
\newcommand{\rr}{\ensuremath{\bf{r}}}

% special values for n
\newcommand{\npart}{\ensuremath{n_{particular}}}
\newcommand{\nliq}{\ensuremath{n_{liquid}}}
\newcommand{\nvap}{\ensuremath{n_{vapor}}}

% Free energy terms
\newcommand{\fid}{\ensuremath{f_\text{ID}(T,n)}}
\newcommand{\fhs}{\ensuremath{f_\text{HS}(T,n)}}
\newcommand{\fdisp}{\ensuremath{f_\text{disp}(T,n)}}
\newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}

% pair potential
\newcommand{\1}{\ensuremath{\textbf{r}_1}}
\newcommand{\2}{\ensuremath{\textbf{r}_2}}
\newcommand{\3}{\ensuremath{\textbf{r}_3}}
\newcommand{\4}{\ensuremath{\textbf{r}_4}}

% fixme
\newcommand{\fixme}[1]{\textcolor{red}{\textbf{[#1]}}}
\newcommand{\davidsays}[1]{\textcolor{blue}{\textit{[#1]}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Applying Renormalization Group Theory to the Square Well Liquid}

\author{Dan Roth}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR
97331}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Near the critical point of a liquid, there are density \fixme{is saying $g(\rr)$ more accurate?} fluctuations at all length scales. In this regime, mean field theories break down, for obvious reasons. As one approaches the critical point, there becomes less and less of a difference in free energy of the liquid and vapor states; this causes difficulty in modeling liquid-vapor coexistence. This work develops a method for calculating liquid-vapor coexistence, and applies renormalization group theory to the square well liquid.

%\tableofcontents
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}\label{sec:theory}

This work applies entirely to the homogeneous case---that is, there are no walls containing our liquid, nor are there any sort of solutes in the liquid. Since our fluid is of infinite extent, the free energy is infinite. Therefore, I work with free energy density:
\begin{align}
  f(T,n) &\equiv \frac{F(T,n)}{V}
\end{align}

\subsection{Renormalization Group Theory}\label{subsec:RGT}
Renormalization group (RG) theory is an iterative procedure that was
developed to deal with large-scale fluctuations of the order parameter
near the critical point of a system. The formulation used here is provided by Forte~\textit{et al}.\cite{Forte11}

\fixme{Discuss Wilson phase cell}

We begin with a baseline free enrgy $f_0(T,n)$. We assume this baseline accounts for all fluctuations of wavelength $\lambda < \lambda_0$. We then double the size of our phase cell and add a correction to the free energy $\delta f_1(T,n)$. This correction accounts for fluctuations with a wavelength $\lambda = 2\lambda_0$. The new free energy is $f_1(T,n) = f_0(T,n) + \delta f_1(T,n)$. Then double the length scale again; $\delta f_2(T,n)$ accounts for fluctionations with $\lambda = 4\lambda_0$ and $f_2(T,n) = f_1(T,n) + \delta f_2(T,n)$. As we continue this process, the $i^{th}$ correction accounts for fluctuations with $\lambda = 2^i\lambda_0$, and $f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)$. In the end, your free energy is
\begin{align}
  f(T,n) = f_0(T,n) + \sum_{i=1}^\infty\delta f_i(T,n) + \fattr
\end{align}

Where $\fattr$ is the contribution due to the longest-wavelenght density fluctuations. At this stage, I do not entirely understand why Forte splits up $\fattr$.

Forte gives $\delta f_i(T,n)$ as
\begin{align}
  \delta f_i(T,n) &= -\frac{\kT}{V_D}\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
\end{align}
where
\begin{align}
  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \bar{f}_D(T,n) + \bar{u}_D(T,n \right) \right] \\
  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\bar{f}_D(T,n) \right]
\end{align}
and
\begin{align}
  \bar{f}_D(T,n) &= \frac{f_{i-1}(T,n+x) + f_{i-1}(T,n-x)}{2} - f_{i-1}(T,n) \\
  \bar{u}_D(T,n) &= \frac{u(T,n+x,\lambda_D) + u(T,n-x,\lambda_D)}{2} - u(T,n,\lambda_D)
\end{align}
Finally, $u(T,n)$ is given by
\begin{align}
  \begin{split}
  u(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \\ &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma
  \end{split}
\end{align}
with
\begin{align}
  \alpha &= \frac{2}{3}\pi\epsilon\sigma^3(\lambda^3 - 1) \\
  \omega^2 &= \frac{1}{5}\sigma^2\frac{\lambda^5 - 1}{\lambda^3 - 1} \\
  \gamma &= \frac{1}{70}\sigma^4\frac{\lambda^7 - 1}{\lambda^3 - 1}
\end{align}

\subsection{Square well liquid free enrgy}\label{subsec:SW}
For the baseline, this work is based on the square well liquid free energy:\cite{Hughes13}
\begin{align}
  f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
\end{align}
The long-range attractive term, $\fattr$ is given by
\begin{align}
  \fattr &= n a_1(n) \ .
\end{align}
$\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere repulsion, and $\fdisp$ is the attraction. The function $a_1(n)$ is the first term from a high-temperature perturbation expansion, described in further detail in sections~\ref{sub2sec:disp} and~\ref{subsec:TPT}. Each function is outlined below.

\subsubsection{Ideal gas}\label{sub2sec:ID}
The ideal gas term is purely local \fixme{whatever that
  means}\davidsays{For your purposes, the purely localness isn't
  really relevant.  It means that when we consider an inhomogeneous
  configuration, the correct answer is just this free energy
  integrated.  i.e. you just add up the local contributions to the
  free energy.  For a homogeneous fluid, the locality isn't really
  relevant.  Except in so far as how it relates to
  fluctuations... since fluctuations are inhomogeneities of a sort.}:
\begin{align}
  \fid &= n\kT\left(\log(n) - 1\right)
\end{align}

\subsubsection{Hard sphere repulsion}\label{sub2sec:HS}
Hard sphere repulsive forces are dealt with by the White Bear version of Fundamental-Mearuse Theory.\cite{Roth02} The hard sphere excess free energy is given in the homogeneous case as
\begin{align}
  \fhs &= \kT\left( \Phi_1(n) + \Phi_2(n) + \Phi_3(n) \right)
\end{align}
The $\Phi_j(n)$ terms are given as
\begin{align}
  \Phi_1(n) &= -n_0\ln(1 - n_3) \\
  \Phi_2(n) &= \frac{n_1n_2}{1 - n_3} \\
  \Phi_3(n) &= n_2^3\left( \frac{n_3 + (1 - n_3)^2\ln(1 - n_3)}{36\pi n_3^2(1 - n_3)^2} \right)
\end{align}
The \emph{fundamental measure densities} are
\begin{align}
  n_0(n) &= \frac{n_2}{4\pi R^2} \nonumber \\
  &= n \\
  n_1(n) &= \frac{n_2}{4\pi R} \nonumber \\
  &= nR \\
  n_2(n) &= n4\pi R^2 \\
  n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
  &= \eta
\end{align}
where $\eta$ is the packing fraction.

\subsubsection{Attraction}\label{sub2sec:disp}
The attractive free energy includes van der Waals attraction between hard spheres and is founded in Thermodynamic Perturbation Theory (TPT).
\begin{align}
  \fdisp &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
\end{align}
The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}.

The first term, $a_1$, is the mean-field dispersion interaction. The second term, $a_2$, describes the effect of fluctuations resulting from compression of the fluid due to the dispersion interaction itself, and is approximated using the local compressibility approximation (LCA), which assumes the energy fluctuation is simply related to the compressibility of a hard-sphere reference fluid.\cite{Barker76}

In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
\textit{monomer} contribution, expressed in terms of energy densities
as:
\begin{align}
  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
\end{align}
with $\alpha^\text{VDW}$ given by
\begin{align}
  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
  \intertext{using reduced units of $x = r/\sigma$, we have}
  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
\end{align}
where $b^\text{VDW}$ is the van der Waals size parameter. It corresponds to the volume excluded by two spherical particles of volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.

Then, the high-temp expansion is an expansion of the monomer term:
\begin{align}
  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
\end{align}

The term $a_1$ is given by
\begin{align}
  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
\end{align}
If we assume random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
\begin{align}
  a_1^\text{VDW} = - n \alpha^\text{VDW},
\end{align}
the van der Waals mean-field energy.

The second-order term is quite a bit more complicated; one must have knowledge of higher-order correlation functions to determine $a_2$. This term describes the response of the attractive energy due to the compression of the fluid from the attractive well. Gil-Villegas \emph{et al.} base their expression on an approximation by Barker and Henderson.\cite{Barker67} Barker and Henderson's approximation considers fluctuations in the number of particles in the potential well. In this way, the fluctuations in $a_2$ are related to the pressure and compressibility of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial n  /\partial P^\text{HS}\right)_T$, we have:
\begin{align}
  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
\end{align}

We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion:\cite{Barker76}
\begin{align}
  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
\end{align}

\subsection{Thermodynamic perturbation theory}\label{subsec:TPT}

\fixme{And introduction to TPT---why do we use it?}\davidsays{We use
  TPT because it gives a good description of simple liquids, as long
  as you are far from the critical point.  Thus it covers most of our
  bases.  It is also an elegant theory, since it comes from a rigorous
  power series expansion in a small parameter... or one that one hopes
  is small. }

This discussion follows primarily from \textit{Theory of Simple Liquids}, by Hansen and McDonald\cite{Hansen06}.

Thermodynamic perturbation theory treats the intermolecular potential as separable into two parts, one for short-range repulsion and another for long-range attraction.

The pair potential $v(\1,\2)$ describes the interaction between two hard spheres. For the square-well model, the pair potential is
\begin{align}
  v_\text{SW}(\1,\2) &=
    \begin{cases}
      \infty & r < \sigma \\
      -\epsilon & \sigma < r < \lambda\sigma \\
      0 & \lambda\sigma < r
    \end{cases}
\end{align}
For this work I use $\lambda = 1.5$.  \davidsays{I think you've gotten
things a little confused.  When $v_\lambda$ is derived below, the $w$
should be the attraction portion, and what appears as $v_{SW}$ should
just be the hard-sphere potential (without attraction).  That's how
the TPT works, it is the attraction that is being added in with the
$\lambda$ proportionality.}

Now, consider this to be the reference, and add a small perturbation:
\begin{align}
  v_\lambda(\1,\2) &= v_\text{SW}(\1,\2) + \lambda w(\1,\2) \label{eqn:small-perturbation}
\end{align}

The total potential energy of our system is
\begin{align}
  V_N(\lambda) = \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j)
  \label{eq:VN}
\end{align}

Given the \emph{configuration integral} (the potential energy component of the partition function)
\begin{align}
  Z_N &= \int d\mathbf{r}^N\, e^{-\beta V_N}
\end{align}
and the non-ideal part of the free energy (i.e.~the excess free energy) is
\begin{align}
  F &= -\kT\ln\left( \frac{Z_N}{V^N} \right)
\end{align}
where $V^N$ is the total volume \fixme{I think?}\davidsays{It is the
  volume taken to the $N^{th}$ power.  It's unfortunate that we have a
  $V_N$ and a $V^N$ that are such different quantities here!}, then
the derivative of the excess free energy with respect to the coupling
parameter $\lambda$ is
\begin{align}
  \frac{\partial F}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
  &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda}
  \label{eqn:dFdlambda} \\
  \hookrightarrow \beta\frac{\partial F}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
  \hookrightarrow \beta F &= \beta F_0 + \beta\int_0^1 d\lambda\, \left\langle V_N^{'}(\lambda) \right\rangle_\lambda
\end{align}
\davidsays{In the above you could simplify things a bit by working out
$dF/d\lambda$ directly (without the $\beta$ factor) which would remove
all the factors of $\beta$.  I would also mention that the subscript
$\lambda$ indicates that the average (and derivative) is evaluated at
$\lambda = \lambda$ (as opposed to $\lambda=0$ which is done later.
Note that at this stage we have an exact equation.  We just took a
derivative of $F$ and then integrated that derivative to find $F$
again.  So far as I can see (and I looked at this derivation a bit
with Eric, who may also be able to help you), $V'_N = \sum w_{ij}$, which
means that $\langle V'_N\rangle = \frac12 N^2 \langle w_{12}
\rangle$, which is potentially a simpler way to look at it.}
where $\left\langle V_N^{'}(\lambda) \right\rangle_\lambda$ is the canonical ensemble average. We can take a series expansion of this average:
\begin{align}\begin{split} % too long; use split environment, and align to a space (&{})
  \left\langle V_N^{'}(\lambda) \right\rangle_\lambda = &{} \left\langle V_N^{'}(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda^2)
\end{split}\end{align}
The derivative term is
\begin{align}
  \frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda} &= \left\langle V_N^{''} \right\rangle_\lambda - \beta\sigma_\lambda^2
\end{align}
where
\begin{align}
  \sigma_\lambda^2 &\equiv \left\langle \left[ V_N^{'}(\lambda) \right]^2 \right\rangle_\lambda - \left\langle V_N^{'}(\lambda) \right\rangle^2_\lambda
\end{align}
\fixme{I don't really understand how this derivative is \ldots
  derived}\davidsays{This is definitely a spot where you can be more
  clear (once you understand).  I believe the $V''_N$ term is zero
  (see above).  Unless I'm severely misunderstanding
  equations~\ref{eqn:small-perturbation} and~\ref{eq:VN}.  The
  derivative is (pretty) easy if you write your ensemble average as a
  multidimensional integral, as in Eq.~\ref{eqn:dFdlambda}.  Write
  down $dF/d\lambda$ in terms of $Z_N$ and an integral of Boltzmann
  factors times $V_N'$.  Then when you take a derivative, it hits top
  and bottom, but each is an integral of an exponential, so those are
  easy to differentiate.  A little math will give you this variance
  formula.} Putting this together, we have
\begin{align}
  \begin{split}
    \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
  \end{split} \\
  \begin{split}
    =&{}  \beta F_0 + \beta\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
  \end{split}
\end{align}
Now, we can define the total perturbation energy as
\begin{align}
  W_N &= \sum_{i=1}^N\sum_{j>i}^N w_\lambda(\mathbf{r}_i,\mathbf{r}_j)
\end{align}
Then, following from equation~\ref{eqn:small-perturbation},
\begin{align}
  V_N^{'} &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \nonumber \\
  &= W_N \\
  \therefore V_N^{''} &= 0
\end{align}
\davidsays{Ah ha! Here you do what I mentioned earlier.  I'd tend to
  put this earlier, so you can throw away the $V_N''$ term sooner.
  And also talking about $W_N$ with this definition I would find nicer
  and clearer than all the $V_N'$.}  With this in mind, we can
simplify equation~\ref{eqn:complicated-HTE} to the
\emph{high-temperature expansion}, first derived by
Zwanzig:\cite{Zwanzig54}
\begin{align}
  \begin{split}
    \beta F = &{} \beta F_0 + \beta\left\langle W_N \right\rangle_0  \\ &{} - \frac{1}{2}\beta^2\left( \left\langle W_N^2 \right\rangle_0 - \left\langle W_N \right\rangle_0^2 \right) + \mathcal{O}(\beta^3)
  \end{split}
\end{align}
Zwanzig further showed that the $n^{th}$ term could be written in terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle W_N\right\rangle_0 \right]^\nu \right\rangle_0$ with $\nu \leq n$.

Writing the statistical averages in terms of the pair density
$n_\lambda^{(2)}(\1,\2)$ (i.e. the probability of having a sphere at
\1 and \2 \fixme{make this more clear?}\davidsays{I'd perhaps phrase
  it as ``probability per volume squared of having one sphere at \1
  and another at \2''}) gives us
\begin{align}
  \frac{\beta F}{N} &= \frac{\beta F_0}{N} + \frac{\beta}{2N}\int_0^1 d\lambda\, \iint d\1 d\2\, n_\lambda^{(2)}(\1,\2)w(\1,\2) \label{eqn:Henderson5-2-12}
\end{align}
\davidsays{Here's another place where all the $N$s on the bottom are
  muddying things up, I think.  (like the $\beta$s were before).  I
  also think this discussion is pretty redundant with parts of the
  previous derivation in terms of $V_N$.}  The pair density itself can
be expanded as
\begin{align}
  n_\lambda^{(2)}(\1,\2) &= n_0^{(2)}(\1,\2) + \lambda\frac{\partial n_\lambda^{(2)}(\1,\2)}{\partial\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2)
\end{align}
Applying this to the \fixme{0th, 1st or 2nd, depending on how you
  think of it}\davidsays{it's ``second'' since ``terms'' are things
  that are added together, while factors are things that are
  multiplied together.  And in Eq~\ref{eqn:Henderson5-2-12} there are
  only two things added together.  (I figured this terminology out on
  my own some time after becoming a professor... it seems to be used,
  but not taught).} term in equation~\ref{eqn:Henderson5-2-12}, we
have
\begin{align}
  \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
  &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)w(\1,\2)
\end{align}
The second-order $\lambda$ term is quite a bit more complicated:
\begin{widetext}
 \begin{align}
   \begin{split}
   \frac{\beta F_2}{N} =&{} -\frac{\beta^2}{2}\left[ \frac{n}{2}\int d\2\, g_0(\1,\2)w^2(\1,\2) \right. \\ &{} + n^2 \iint d\2 d\3\, g_0^{(3)}(\1,\2,\3)w(\1,\2)w(\1,\3) \\ &{} \left. + \frac{n^3}{4}\iiint d\2 d\3 d\4\, \left( g_0^{(4)}(\1,\2,\3,\4) - g_0^{(2)}(\1,\2)g_0^{(2)}(\3,\4) \right)w(\1,\2)2(\3,\4) \right] \\ &{} - \frac{1}{4}S_0(0)\left[ \frac{\partial}{\partial n}\left( n^2\int d\2\, g_0(\1,\2)w(\1,\2) \right)^2 \right]
   \end{split}
 \end{align}
\end{widetext}
This term requires use to know the form of higher-order correlation functions and the structure function $S_0(k)$ of the reference system. These things are very difficult. Hence, we use approximations, as outlined in section~\ref{sub2sec:disp}.

\davidsays{I'm thinking this whole section should go \emph{before} the
  section on the square well free energy, since that section uses this
  method (in approximation).  Then you can say something to the effect
  of ``in the next section we will see a simple approximation for this
  term.''  Hansen's explanation of TPT is not very clear, which makes
  it all the more valuable to make \emph{this} a clear and lucid
  explanation.  I'd love to see just a couple more steps at key
  places---largely where you were confused.  Probably it's best to
  actually talk about this section in person before rewriting.  This
  has been an excellent first shot, though! }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}\label{sec:methods}

\subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
Coexistence curves are found by modifying the chemical potential $\mu$
in the grand free energy per volume $\phi(T,n)/V$. Grand free energy
density is defined as
\begin{align}
  \frac{\Phi(T,n)}{V} &= \phi(T,n) \nonumber \\
                 &= f(T,n) - n\mu \nonumber \\
                 &= f(T,n) - n\frac{df(T,n)}{dn}\bigg|_{\npart}\ .
\end{align}
It is the value $\npart$ that we adjust to find coexistence curves.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
  \caption{Grand free energy per unit volume, SW, low temp}
  \label{fig:SW-phi-lowT}
\end{figure}

The grand free energy density generally has two distinct minima when
plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an example.) The liquid will be in liquid-vapor equilibrium
when those minima have the same value for $\phi(T,n)$. There is a local
maximum between those minima, and the value of the density at that
maximum is used for $\npart$. Find this value at a given temperature
then plot $\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new
value for $\npart$ such that the two minima have the same value for
$\phi(T,n)$. (It is helpful to note that if $\npart$ is increased,
$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$ is
decresed.) The program I wrote uses the previous
temperature's $\npart$ as a ``first guess,'' maximizing
$\phi(T,\npart)$. From there, the program adjusts $\npart$
until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
uses this new value of $\npart$ as the ``first guess'' at $\npart$ for the next
higher temperature.

The algorithm is as follows; start at some temperature $T=T_{start}$, with an intial guess for $\npart$.
\begin{enumerate}
  \item Calculate $\nvap$ and $\nliq$ by calculating the minima in $\phi(T_{start},n)$ over appropriate regions
  \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
  \item Calculate \[\delta\mu = \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\nliq - \nvap}\]
  \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + \delta\mu$
  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
  \item Calculate $\phi(T_{start},\npart^*)$
  \item Re-caluclate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some computational tolerance (I used $tol=10^{-5}$)
  \item Increase the temperature to $T = T + \delta T$ and repeat
\end{enumerate}

\subsection{Free Energies}\label{subsec:free-energies}
The $a_j(n)$ terms require even more simplification when
``converting'' from the mathematical formulation to Python code, to
rid ourselves of troublesome integrals. Gil-Villegas gives us
\newcommand\eff{\textit{eff}}
\begin{align}
  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{\eff}) \\
  &= -4\eta\epsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{\eff}/2 \right)}{(1 - \eta_{\eff})^3}
\end{align}
and
\begin{align}
  \eta_{\eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
\end{align}
with
\begin{align}
  \left( \begin{array}{c}
    c_1 \\
    c_2 \\
    c_3
    \end{array} \right)
  &= \left( \begin{array}{ccc}
    2.25855 & -1.50349 & 0.249434 \\
    -0.669270 & 1.40049 & -0.827739 \\
    10.1576 & -15.0427 & 5.30827
    \end{array} \right)
  \left( \begin{array}{c}
    1 \\
    \lambda \\
    \lambda^2
    \end{array} \right)
\end{align}

\subsection{RG partition functions}\label{subsec:fbar-ubar}
I use the midpoint method for evaluating the integrals of $\bar{f}_D$ and $\bar{u}_D$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and discussion}

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
  \end{center}
  \caption{Liquid-vapor coexistance for a square well liquid.}
  \label{fig:coexistance_SW}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
  \end{center}
  \caption{Free energy density for RG at low temperature}
  \label{fig:phi-RG-lowT}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
  \end{center}
  \caption{Free energy density for RG near critical temp}
  \label{fig:phi-RG-highT}
\end{figure}

Figures~\ref{fig:phi-RG-lowT} and~\ref{fig:phi-RG-highT} show the grand free energy per unit volume at a low temperature, and near the critical point. It is clear that the program is not de-bugged.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}
I have implemented a the use of renormalization group theory in calculations for the square well liquid. The program is not entirely de-bugged. I have also developed an algorithm that calculates liqiud-vapor coexistence curves, even close to the critical point.

\bibliographystyle{unsrt}
\bibliography{project} % Produces the bibliography via BibTeX.

\end{document}