### 2015-04-28 papers/histogram/thesis/thesis.tex

\documentclass[11pt]{article}

%%% standard header
\usepackage[margin=1in]{geometry} % one inch margins
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancyplain} % page formatting style
\usepackage{hyperref} % for linking references
\newcommand{\psl}{6pt} % store spacing as command for consistency
\setlength{\parindent}{0cm} % don't indent new paragraphs...
\parskip \psl % ... place a space between paragraphs instead
\usepackage[inline]{enumitem} % include for \setlist{}, use below
\frenchspacing % add a single space after a period
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% physics
\usepackage[boldvectors,braket]{physymb} % physics package
\newcommand{\bk}{\Braket} % shorthand for braket notation

%%% math
\usepackage{amsmath} % math package
\renewcommand{\t}{\text} % for text in math environment
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis

%%% figures
\usepackage{graphicx,float,subcaption} % floats, etc.
\usepackage{footnote} % for footnotes in floats
\usepackage[font=small,labelfont=bf]{caption} % caption text options

%%% bibliography, table of contents
\usepackage[sort&compress,numbers]{natbib} % bibliography options
\bibliographystyle{apsrev4-1}
\usepackage[english]{babel} % allow editing bibliography title
\addto\captionsenglish{ % set bibliography title
  \renewcommand{\contentsname}{Table of Contents}
  \renewcommand\refname{\bf References}}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

%%% algorithm float
\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}
{\hrulefill\begin{enumerate}}
{\end{enumerate}\hrulefill}

\renewcommand{\title}{Optimizing Monte Carlo state space exploration
  of the square-well fluid}
\renewcommand{\author}{Michael A. Perlin}
\newcommand{\institution}{Oregon State University}
\newcommand{\department}{Department of Physics}
\newcommand{\supervisor}{Dr. David Roundy}
\renewcommand{\date}{21 April 2015}

%%% notes, etc.
\usepackage{color}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\fixme}[1]{[\red{fixme:} \emph{#1}]}

\fancypagestyle{abstract}{
  \rhead{}
  \lhead{}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

  \center

  \textsc{\LARGE \institution}\\[1.5cm]
  \textsc{\Large \department}\\[0.5cm]

  \HRule \\[0.4cm]
  { \huge \bfseries \title}\\[0.4cm]
  \HRule \\[1.5cm]

  \begin{minipage}{0.4\textwidth}
    \begin{flushleft} \large
      \emph{Author:}\\
      \author
    \end{flushleft}
  \end{minipage}
  ~
  \begin{minipage}{0.4\textwidth}
    \begin{flushright} \large
      \emph{Supervisor:} \\
      \supervisor
    \end{flushright}
  \end{minipage}\\[4cm]

  {\large\date}

  \vfill

\end{titlepage}

\thispagestyle{empty}

\newpage

\tableofcontents

\thispagestyle{empty}

\newpage

\listof{algorithm}{\Large List of Algorithms}

\listof{figure}{\Large List of Figures}

\thispagestyle{empty}

\newpage

\setcounter{page}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
\label{sec:abstract}

\thispagestyle{abstract}

We wish to identify and develop efficient numerical methods for
determining thermodynamic properties of the square-well fluid in order
to test square-well density functional theories. The liquid-vapor
phase interface is an interesting regime for testing density
functional theories, but unbiased (i.e. Metropolis) Monte Carlo
simulations are incapable of sampling the low energy fluid states
which dominate the partition function at low (i.e. liquid-state)
temperatures. Previous works have developed several generic Monte
Carlo histogram methods for collecting statistics on low energy system
states, but little or no literature exists on these methods'
systematic comparison, as well as their application to the square-well
fluid. The square-well fluid in particular introduces application
challenges not manifest in traditional models for testing and
benchmarking such numerical techniques (e.g. the Ising model). We
propose to implement these methods in Monte Carlo simulations of the
square-well fluid, determine appropriate performance metrics for their
comparison, and identify those methods which are most efficient and
effective at sampling otherwise improbable system states.

\fixme{Rewrite after the rest of the thesis is completed}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

\fixme{give more intro about phase transitions and the critical
  point?}

Understanding the behavior of fluids near the critical point, or the
point at which the distinction between liquid and gas ceases to become
well-defined, is a major challenge in the field of classical density
functional theory. Studying the critical point of a fluid via
simulations requires exploring the boundary between its liquid and
gaseous phases. The square-well fluid is the simplest system with a
liquid-vapor phase transition, and is therefore of interest for
studying the critical point of fluids. For a generic square-well fluid
to be observed in a liquid state, it must have a localized
distribution of the spheres which make up the fluid. Such a
distribution implies a low energy fluid state, as all interactions
between spheres are attractive. Conversely, a low energy liquid state
of a square-well fluid otherwise observable in both liquid and gaseous
phases implies a localized distribution of spheres. The search for
liquid states of the square-well fluid is thus equivalent to the
search for low energy states.

Monte Carlo simulations are a standard means for studying the
equilibrium thermodynamic properties of a system. In the most
straightforward implementation of Monte Carlo fluid simulations,
however, highly localized distributions of fluid elements are
extremely unlikely to occur, making such simulations impractical for
studying the critical point of the square-well fluid. Several generic
computational methods, called histogram methods, exist for dealing
with such a problem; namely, that interesting regions of a system's
state space cannot be sampled via standard Monte Carlo methods in a
reasonable amount of time. There is no common consensus, however, on
the relative efficacy of these methods. Furthermore, some methods,
such as the Wang-Landau method\cite{wang_landau}, require tuning
several free parameters that lack both ``obvious'' values and
systematic means for assigning them. This work will thus attempt to
quantify and compare the effectiveness of a few such methods as
applied to the square-well fluid.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:background}

\subsection{The square-well fluid}
\label{sec:sw_fluid}

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.47\textwidth]{figs/square-well.pdf}
  \caption[The square-well pair potential]{The square-well pair
    potential can be used to model short range forces to first
    order. The infinite potential at $r<\sigma$ simply enforces the
    condition that spheres cannot overlap. At $r>\lambda\sigma$, the
    potential is null, and makes no contribution to the fluid's
    internal energy.}
  \label{fig:pair_potential}
\end{figure}

The square-well (SW) fluid is a simple model used in classical density
functional theories to capture low order effects of short-range
attractive forces, such as the van der Waals force. The fluid is
composed of spheres with diameter $\sigma$ which have a pair potential
\begin{align}
  v_{sw}\p{\vec r}=\left\{
    \begin{array}{ll}
      \infty & \abs{\vec r}<\sigma \\
      -\epsilon & \sigma<\abs{\vec r}<\lambda\sigma \\
      0 & \abs{\vec r}>\lambda\sigma
    \end{array}
  \right., \label{eq:pair_potential}
\end{align}
also shown graphically in Figure \ref{fig:pair_potential}, where the
parameters $\lambda$ and $\epsilon$ are referred to as the well width
and depth, respectively. The first ($\abs{\vec r}<\sigma$) part of
this potential forbids spheres from overlapping, whereas the second
($\sigma<\abs{\vec r}<\lambda\sigma$) associates an energy $-\epsilon$
with each pair of spheres whose centers are within distance of
$\lambda\sigma$ of each other (where typically, $\lambda\in(1,3]$).
The net potential energy of the square-well fluid is thus
\begin{align}
  E=\sum_{i<j}v_{sw}\p{\vec r_i-\vec r_j},
  \label{eq:internal_energy}
\end{align}
where $\vec r_i$ is the position of the $i$-th sphere. As the
potential energy $E$ is the primary form of energy concerning us in
this paper, we will refer to $E$ as simply the ``energy'' of the
fluid. An important feature of the square-well fluid is that its
energy is always an integer multiple of the well depth. A homogeneous
square-well fluid is thus uniquely identified by its well width
$\lambda$ and filling fraction $\eta$ (i.e. the proportion of space
filled by spheres; a dimensionless density), as all other properties
can be normalized to the natural energy scale $\epsilon$ and length
scale $\sigma$. In practice, our simulation codes use dimensionless
energies $E/\epsilon$, temperatures $kT/\epsilon$, and distances
$r/\sigma$.

\subsection{Monte Carlo fluid simulations}
\label{sec:mc_sim}

While model systems are powerful tools for understanding complex
physical systems, they do not themselves exist in the real
world. Consequently, direct experimental tests of theories for model
system (e.g. square-well density functional theories) are not
possible. For this reason, model system theories are commonly tested
against Monte Carlo simulations. Proper implementation of Monte Carlo
methods to study completely characterized systems ensures that
statistical results from simulations converge on the exact properties
of the simulated system in the infinite simulation time
limit. Furthermore, uncertainties in quantities computed via Monte
Carlo simulations are typically well-defined, monotonically decreasing
functions of simulation time, allowing one to run simulations to the
desired level of accuracy.

\subsubsection{Implementation}
\label{sec:mc_implementation}

\begin{algorithm}[tb]
  \caption{Metropolis Monte Carlo fluid simulation}
  \label{alg:MC}
  \begin{alg}

  \item Construct an initial ``typical'' (i.e. non-ordered) fluid
    configuration. \fixme{explain how}

  \item Attempt to change the position of one sphere (in general, a
    single fluid ``atom'' or ``molecule'') to a random location,
    rejecting the change if it results in a forbidden fluid
    configuration (e.g. two or more spheres overlap) and accepting the
    change otherwise. Attempting to move a single sphere is referred
    to as a move. \label{alg:mc_move}

  \item Repeat step \ref{alg:mc_move} for every other sphere in the
    fluid. Attempting to move every sphere once is referred to as an
    iteration of the simulation.
    \label{alg:mc_iteration}

  \item Repeat step \ref{alg:mc_iteration} indefinitely, or until data
    of sufficient quality has been generated, periodically collecting
    and dumping statistics on fluid states (e.g. energy, pair
    distribution histograms, etc.) to data files.

  \end{alg}
\end{algorithm}

Algorithm \ref{alg:MC} provides a sketch of unbiased, Metropolis Monte
Carlo fluid simulations. Such an algorithm is ``unbiased'' in the
sense that it collects statistics (i.e. data) on all valid system
configurations with equal probability. Statistics whose collection
time scales as $\mathcal O\p{1}$ with system size, meaning that
increasing the number of spheres $N$ in the simulation does not affect
collection time, can be collected after every move (defined in the
algorithm). Statistics whose collection time scales as $\mathcal
O\p{N}$, meaning that doubling $N$ doubles collection time, can be
collected after each iteration. In general, collection with $\mathcal
O\sp{\chi\p{N}}$ time scaling should not occur more often than once
every $\chi\p{N}$ moves (where $\chi\p{N}$ may be an arbitrary
function of $N$ void of constant prefactors, e.g. $N\log N$, or
$2^N$). These collection rules ensure that scaling up simulations does
not cause them to asymptotically spend all computation time collecting
statistics, and no time actually simulating the fluid. Collected
statistics are used to find thermodynamic properties of the simulated
fluid.

In this work, we are concerned with simulating the \emph{homogeneous}
square-well fluid. To avoid edge effects resulting from fluid behavior
near a wall, we employ periodic boundary conditions. The use of a
finite cell with periodic boundary conditions suppresses all density
fluctuations on scales larger than the dimensions of the simulated
fluid cell, thereby introducing a source of error. Addressing and
sequestering this error, however, is outside the scope of this work,
and involves considering the limit of numerical results as the number
of spheres $N\to\infty$ (keeping the filling fraction $\eta$
constant).

\subsubsection{Computing observables}
\label{sec:computing_observables}

The sort of Monte Carlo simulations described above are only capable
of collecting statistics on functions of system microstates, and
finding correlations between these functions. For example, in
simulation one might periodically compute both the energy $E\p{s}$ and
some system property $X\p{s}$, both of which are determined by the
microstate $s$, in order to find the mean value of $X$ at any given
energy $E$, i.e. $\bk{X}_E$. In the real world, however, information
about a system's microstates, and by extension functions of
microstates (e.g. the energy $E$), is inaccessible. One therefore
cannot measure $\bk{X}_E$ directly. Instead, one typically measures
the dependence of thermodynamic properties on macroscopic state
variables such as temperature, i.e. $\bk{X}_T$.

To find $\bk{X}_T$, we first need to understand the concept of a
density of states. Given an arbitrary system property $Y\p{s}$
(e.g. energy) determined by the microstate $s$, one can define a
density of states $D\p{Y}$ such that
\begin{enumerate*}[label=\roman*)]
\item for arbitrary $Y_0$, the value of $D\p{Y_0}$ is proportional to
  the number of microstates $s$ for which $Y\p{s}=Y_0$, and
\item $\sum_YD\p{Y}=1$.
\end{enumerate*}
The units of $D\p{Y}$ are inverse to the units of $Y$. The temperature
dependence of $\bk{X}_T$ can be expressed in terms of the expectation
value $\bk{X}_E$ and the density of states $D\p{E}$ by
\begin{align}
  \bk{X}_T=\f1{Z\p{T}}\sum_E\bk{X}_ED\p{E}e^{-E/kT},
  \label{eq:XT_norm}
\end{align}
where the partition function $Z\p{T}$ is simply a normalization
factor, given by
\begin{align}
  Z\p{T}=\sum_ED\p{E}e^{-E/kT}.
\end{align}
To reduce redundant computations and numerical error in
implementation, we will generally use a partition function with an
unnormalized density of states $\tilde D\p{E}$, i.e.
\begin{align}
  \tilde Z\p{T}=\sum_E\tilde D\p{E}e^{-E/kT},
\end{align}
in terms of which
\begin{align}
  \bk{X}_T=\f1{\tilde Z\p{T}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT}.
  \label{eq:XT}
\end{align}

\subsection{Histogram methods}
\label{sec:histogram_methods}

Due to the fact that Metropolis Monte Carlo simulations sample all of
state space randomly and without preference, a histogram $H\p{X}$ of
observations of some system property $X$ is directly proportional to
the density of states $D\p{X}$; that is, the number of observations
$H\p{X_0}$ of the system with $X\p{s}=X_0$ is proportional to the
total number of states $s$ for which $X\p{s}=X_0$. It is sometimes the
case, however, that the density of states in some region $R$ in the
range of possible $X$ is so low that it is practically impossible (via
Metropolis Monte Carlo) to sufficiently sample $R$, that is,
accumulate a statistically significant histogram $H\p{X\in R}$, in any
reasonable amount of time.

\subsubsection{Biased sampling}
\label{sec:biased_sampling}

Histogram methods provide a means to address Metropolis Monte Carlo's
inability to sufficiently sample regions of parameter space with a low
density of states by introducing a bias into the otherwise random
sampling of state space. A weighting function $w$ is introduced, whose
domain is the value of some property $X\p{s}$ which depends on the
system state $s$. An additional condition is then added to step
\ref{alg:mc_move} of Algorithm \ref{alg:MC} in order to accept an
attempted move: the weights $w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of
the initial (pre-move) state $s_i$ and final (post-move) state $s_f$
are used to determine the probability $P\p{s_i\to s_f}$ of accepting
an otherwise valid move via
\begin{align}
  P\p{s_i\to s_f}=\min\set{\f{w\sp{X\p{s_f}}}{w\sp{X\p{s_i}}},1}.
  \label{eq:move_prob}
\end{align}
This formula means that when $w\sp{X\p{s_f}}>w\sp{X\p{s_i}}$, the move
$s_i\to s_f$ is accepted; when $w\sp{X\p{s_f}}<w\sp{X\p{s_i}}$, the
ratio of these weights determines the probability of accepting the
move. Due to the fact that only ratios of weights determine
$P\p{s_i\to s_f}$, the weights $w\sp{X\p{s}}$ are scale-invariant,
meaning that their effect on simulations is unchanged by scale factors
that are constant with respect to $X\p{s}$.

\begin{algorithm}[tb]
  \caption{Biased Monte Carlo fluid simulation}
  \label{alg:biased_MC}
  \begin{alg}

  \item Construct an appropriate weight function $w\sp{X\p{s}}$, whose
    argument is a system property $X\p{s}$ determined by the
    microstate $s$.

  \item Construct an initial typical fluid configuration.

  \item Attempt to change the position of one sphere, rejecting the
    change if
    \begin{enumerate*}[label=\roman*)]
    \item it results in a forbidden fluid configuration, or
    \item a newly chosen random number on the interval $\sp{0,1}$ is
      larger than the probability given by (\ref{eq:move_prob}).
    \end{enumerate*}
    \label{alg:biased_mc_move}

  \item Repeat step \ref{alg:biased_mc_move} for every other sphere in
    the fluid.
    \label{alg:biased_mc_iteration}

  \item Repeat step \ref{alg:biased_mc_iteration} until data of
    sufficient quality has been generated.

  \end{alg}
\end{algorithm}

Employing biased Monte Carlo simulations, sketched out in Algorithm
\ref{alg:biased_MC}, allows one to construct weight functions that
favor some region of state space over others, as transitions to states
with higher weights are always accepted, whereas transitions to states
with lower weights may be rejected, artificially preventing the
simulated system from leaving interesting regions of state
space. Crucially, the bias introduced by weights can be reversed when
computing system properties from sampling statistics, as the weight of
a particular state is directly proportional to the probability bias of
that state; that is, a state with a weight of 2 will be sampled twice
as often as it would have been with a weight of 1. If we wish to
convert a histogram $H\p{X}$ of observations in a biased Monte Carlo
simulation into a numerical unnormalized density of states $\tilde
D\p{X}$ in $X$, we therefore divide the histogram by the corresponding
weights $w\p{X}$, i.e.
\begin{align}
  \tilde D\p{X}=\f{H\p{X}}{w\p{X}}.
  \label{eq:dos}
\end{align}
The normalized density of states $D\p{X}$ in $X$ is then
\begin{align}
  D\p{X}=\f{\tilde D\p{X}}{\sum_X\tilde D\p{X}}
  =\f{H\p{X}/w\p{X}}{\sum_XH\p{X}/w\p{X}}.
  \label{eq:dos_norm}
\end{align}
In general, the domain and shape of the weight function will depend on
the desired yields (e.g.  density of states, heat capacity) of a
simulation. A histogram method is simply an algorithm or procedure for
determining a weight function appropriate for a particular simulation.

In this paper, we will consider weights $w\p{E}$ which depend only on
the energy $E$ of the system. Due to the fact that the square-well
fluid can only have discrete energies $E=-n\epsilon$, where $n$ is a
non-negative integer and $\epsilon$ is the well depth, in simulation
we store the weight function as an array of values. We may therefore
use the terms ``weight function,'' and ``weight array,'' and
``weights'' interchangeably.

\fixme{mention that simulating with flat weights $w\p{E}=w_0$ is the
  same as running a Metropolis MC simulation.}

\subsubsection{Canonical weights}
\label{sec:canonical_weights}

The most common weight array $w\p{E}$ used by physicists for what is
called a canonical Monte Carlo simulation involves choosing a
particular temperature $T_0$, and using weights proportional to the
Boltzmann factor at that temperature, i.e.
\begin{align}
  w\p{E}=e^{-E/kT_0},
\end{align}
where there is no reason to normalize $w\p{E}$ due to the fact that
only ratios of weights, as per (\ref{eq:move_prob}), are ever used in
simulation. The fact that only ratios of weights are used in
simulation makes $w\p{E}$ scale-invariant.

The partition function at $T=T_0$ for simulations with canonical
weights is
\begin{align}
  \tilde Z\p{T_0}=\sum_E\tilde D\p{E}e^{-E/kT_0}
  =\sum_E\f{H\p{E}}{w\p{E}}~e^{-E/kT_0}
  =\sum_E\f{H\p{E}}{e^{-E/kT_0}}~e^{-E/kT_0}=\sum_EH\p{E},
  \label{Z_canonical}
\end{align}
and the value of a thermodynamic property $X\p{T_0}$
\begin{align}
  X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT_0}
  =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}}.
  \label{eq:X_canonical}
\end{align}
The simplifications in in (\ref{Z_canonical}) and
(\ref{eq:X_canonical}), which have no explicit dependence on $T_0$,
occur because canonical Monte Carlo simulations sample energies in
proportion to the distribution (over energy) of microstates at a
temperature of $T_0$.  Canonical Monte Carlo simulations can therefore
fail to sufficiently sample energies which are important
(i.e. energies with a non-negligible state probability density) at
different temperatures. As a consequence, such simulations should not
be used to determine properties $X\p{T\ne T_0}$, and are thus referred
to as ``fixed temperature'' simulations.

Though canonical Monte Carlo is simple to implement, its inability to
investigate a system at more than one temperature at a time is a
disadvantage for determining the temperature dependence of system
properties. In order to find the behavior of $\bk{X}_T$, one must run
many simulations at discrete temperature intervals; each such
simulation will yield one sample (e.g. one data point) of $\bk{X}_T$.

Sampling low energies, however, or understanding system behavior at
low temperatures, is even more problematic with canonical
weights. Using low temperature canonical weights will indeed force a
simulated system down to low energies, but will also likely freeze the
system into a local minimum of its energy landscape. Freezing into a
state means that a simulation will sample only a small portion of the
energy landscape, even though there may (and generally will) be many
other states with the same energy.

Due to these problems, we will not use canonical weights alone to
study the square-well fluid. We will, however, use canonical weights
for part of all weight arrays, as discussed in Section
\ref{sec:min_energy}.

\subsubsection{Broad energy sampling}
\label{sec:broad_energy_sampling}

Given the expression for $\bk{X}_T$ in (\ref{eq:XT}), sufficient
accumulation of statistics on $\bk{X}_E$ at all available energies in
principle allows one to determine $\bk{X}_T$ for any temperature
$T$. In practice, the density of states can fall off so quickly with
energy that some range of allowable energies is practically
inaccessible via Monte Carlo simulations, biased or otherwise. In such
a case, computing $\bk{X}_T$ to a reasonable degree of accuracy
requires sufficiently sampling the energies at which $\tilde
D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:XT}). Section
\ref{sec:energy_range} provides more discussion on this subject,
namely on identifying the energies which are important for computing
system properties at a given temperature.

We will employ histogram methods in order to sample as broad of an
energy range as possible in each simulation. Broad energy sampling
will allow us to determine, for various square-well fluids,
\begin{enumerate*}[label=\roman*)]
\item the density of states $D\p{E}$, and
\item the temperature dependence $\bk{X}_T$ of various thermodynamic
  properties $X$, particularly at temperatures near the liquid-gas
  phase boundary.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\label{sec:methods}

\fixme{Give an overview of the methods section}

\fixme{Make a note on storing the logarithm of weights and
  periodically rescaling them (resetting the zero on $\ln w$) to
  reduce numerical error.}

\subsection{The flat histogram (multi-canonical) method}
\label{sec:flat_histogram}

The flat histogram method, also called the multi-canonical method,
assumes complete knowledge of the density of states $D\p{E}$ of the
system in question, and solves (\ref{eq:dos}) for a weight array
$w\p{E}$ which should yield a flat energy histogram $H\p E=H_0$ to get
\begin{align}
  w\p E=\f1{D\p E},
  \label{eq:flat_weights}
\end{align}
where we may neglect the constant scale factor $H_0$ and do not worry
about normalization factor distinguishing $D\p{E}$ from $\tilde
D\p{E}$ due to the scale-invariance of $w\p{E}$.

For all but the most trivial or well-studied systems, however, the
density of states $D\p{E}$ is not known prior to simulation, and is in
fact one of the system properties which a simulation is intended to
determine. Nonetheless, the this method has pedagogical value by
enabling discussion of an ``optimal'' weight array for generating a
flat energy histogram in simulations. All algorithms to construct
$w\p{E}$ with the aim of producing a flat energy histogram should
converge on the same weight array, given by (\ref{eq:flat_weights}).

\begin{algorithm}[b]
  \caption{A naive flat histogram method}
  \label{alg:flat_histogram}
  \begin{alg}

  \item Construct an initial typical fluid configuration.

  \item Simulate (without weights) for some time $\tau$, which may be
    real time, computer time, or a number of simulation iterations,
    collecting a histogram $H\p{E}$ of energy observations after every
    move.

  \item Set weights $w\p{E}=1/H\p{E}$. At energies with $H\p{E}=0$,
    set $w\p{E}=1$.

  \end{alg}
\end{algorithm}

Algorithm \ref{alg:flat_histogram} provides a naive way to implement
the flat histogram method without knowing a priori the density of
states $D\p{E}$. This algorithm involves first running a simulation
without weights (or with constant weights $w\p{E}=w_0$) while
collecting a histogram $H\p{E}$ of the observed energies $E$. After
some time (loosely defined), the histogram $H\p{E}$ may be taken as an
approximation of the density of states $\tilde D\p{E}$ (see Section
\ref{sec:histogram_methods}), and used via (\ref{eq:flat_weights}) to
generate a weight array for an actual simulation. Algorithm
\ref{alg:flat_histogram} has one free parameter, which determines the
amount of time for which to collect an energy histogram $H\p{E}$.

The problem with this implementation of the flat histogram method is
that histogram methods are generally necessary when some range of
energies $R$ is inaccessible via Metropolis Monte Carlo
simulations. In such a case, the energy histogram $H\p{E\in R}$ after
simulating for any reasonable amount of time will be statistically
insignificant or null, invalidating the approximation
$H\p{E}\approx\tilde D\p{E}$. This method will therefore only
``flatten'' the histogram at energies $E$ which have been sufficiently
sampled for $H\p{E}$ to be statistically significant. The end goal in
implementing histogram methods, however, is precisely to sample those
energies that cannot be sufficiently sampled without employing clever
schemes to bias Monte Carlo simulations. The flat histogram method
therefore does not directly help us study the square-well fluid, but
it does provide the basis for the method in Section
\ref{sec:simple_flat}.

\subsection{The simple flat method}
\label{sec:simple_flat}

The ``simple flat'' method, developed by ourselves, extends the
implementation of the flat histogram method discussed in Section
\ref{sec:flat_histogram} to an algorithm intended to be the simplest
viable broad energy histogram method. While we do not expect this
method to outperform other, more sophisticated methods, the simple
flat method should provide a lower bar and a standard of comparison
for the performance of other methods. Any nontrivial histogram method
should, at a minimum, consistently outperform the simple flat method
in order to be considered for any applications.

The idea behind the simple flat method is such: after simulating for
some time, the energy histogram $H\p{E}$ will have a peak, or some
other uneven distribution of observations, in a region of energies $R$
for which $H\p{E\in R}$ statistically significant. One can thus use
the existing histogram to construct weights which should flatten
$H\p{R}$. Subsequent simulations should then spend less time at
energies with the highest state densities, and more time at energies
with relatively low state densities. The energy histogram from such
simulations should therefore be statistically significant in a larger
range of energies, improving the available estimate for a density of
states with which to compute flat histogram weights. One could then
use the improved estimate to construct new weights, and repeat this
simulation and updating process until simulations satisfy some end
condition (in fact, most histogram methods will require end
conditions, which is discussed in Section \ref{sec:end_conditions}).

\begin{algorithm}[!t]
  \caption{The simple flat method}
  \label{alg:simple_flat}
  \begin{alg}

  \item Construct an initial typical fluid configuration.

  \item Set $k=0$, choose an initial number of iterations $n$ for
    which to simulate, and initialize weights $w\p{E}=1$ for all $E$.

  \item Simulate for $n$ iterations (with weights), collecting a
    histogram $H\p{E}$ of energy observations after every move.
    \label{alg:simple_flat_sim}

  \item If some predetermined initialization end condition has
    \emph{not} been met, then
    \begin{enumerate}
    \item increment $k\leftarrow k+1$
    \item set $n_k=un_{k-1}$ with a predetermined factor $u\ge1$,
      \label{alg:simple_flat_f}
    \item update weights as $w\p{E}\leftarrow w\p{E}/H\p{E}$ for all
      $E$ at which $H\p{E}\ne0$,
      \label{alg:simple_flat_weight_update}
    \item return to step \ref{alg:simple_flat_sim}.
    \end{enumerate}

  \end{alg}
\end{algorithm}

Algorithm \ref{alg:simple_flat} provides an implementation of the
simple flat method. This algorithm has two free parameters: the
initial number of iterations $n_0$ for which to simulate and the
factor $u$. Due to the exponential growth of $n_k=u^kn_0$, this
algorithm should not be sensitive to the value of $n_0$.

The most interesting part of Algorithm \ref{alg:simple_flat} is step
\ref{alg:simple_flat_weight_update}, which updates the weights as
$w\p{E}\leftarrow w\p{E}/H\p{E}$. The reasoning behind this step is
such: say that for two energies $E_0$ and $E_b$ we have that
$H\p{E_0}=\bk{H}_E$ and $H\p{E_b}=b\bk{H}_E$, where $\bk{H}_E$ is the
mean histogram value over all energies. In this case, we wish to
decrease the current simulation bias on $E_b$ by a factor of $b$ while
keeping the bias on $E_0$ the same. As the weights $w\p{E}$ are
proportional to the simulation bias on each respective energy $E$,
dividing the weights $w\p{E}$ by the current histogram $H\p{E}$
achieves the desired changes to the simulation biases. The factor
$\bk{H}_E$ in this weight update has no effect on the function of the
weights, and is in practice immediately scaled out of the weight
array.

In our implementation of the simple flat method, we used $u=2$ and
$n_0=L\exp\p{\epsilon/kT_{\t{min}}}$, where $L$ is the number of
energy levels of the given square-well fluid and $T_{\t{min}}$ is a
minimum temperature of interest for the current simulation. The factor
of $L$ appears because initialization time should scale with the range
of energies of the simulated system. If we have twice the number of
energies, we should explore energy space for twice as long. To
estimate $L$, we multiplied the maximum number of spheres $M$ which
fit within a radius of $\lambda\sigma$ of a single sphere on a
face-centered cubic lattice (i.e. the maximum number of spheres with
which a given sphere could interact) by the total number of spheres
$N$, and divided the result by two so as to not double-count
interactions between spheres (so $L=MN/2$). The exponential
$\exp\p{\epsilon/kT_{\t{min}}}$, a Boltzmann factor for two adjacent
energies, appears because lower minimum temperatures of interest
warrant sampling lower energies, which in turn requires simulating for
longer (in step \ref{alg:simple_flat_sim} of Algorithm
\ref{alg:simple_flat}) because these energies have lower state
densities. The introduction of a minimum temperature may seem ad-hoc,
but all simulations in this paper require a choice of minimum
temperature anyways (addressed in Section \ref{sec:energy_range}). In
this paper, we will always use $kT_{\t{min}}=0.2\epsilon$.

\subsection{The Wang-Landau (WL) method}
\label{sec:wang_landau}

The Wang-Landau method is the first published histogram method we
discuss\cite{wang_landau, wang_landau_analysis, wang_landau_mod}. This
method has been used to study a wide variety of systems\fixme{which
  systems? cite}, and is the standard method\fixme{cite} to achieve
broad energy sampling via Monte Carlo simulations. \fixme{more intro
  on WL}

Unlike the simple flat method, the Wang-Landau method modifies the
weight array on the fly. After every move during initialization
Wang-Landau decreases the weight $w\p{E}$ on the current energy $E$ by
some factor of $f$, so as to decrease weights on the most commonly
observed energies relative to those of the least commonly observed
energies. Eventually, such a simulation with constant modification of
the weights $w\p{E}$ should yield a flat histogram
$H\p{E}\approx\bk{H}_E$ for all $E$, as the modifications to $w\p{E}$
increasingly pushes the simulated system away from the most sampled
energies, and toward the least sampled energies. When the energy
histogram is sufficiently flat, $H\p{E}$ is reset (i.e. set to
$H\p{E}=0$ for all $E$), the factor $f$ is decreased (geometrically
approaching unity), and the entire process is repeated until $f-1$
falls below some cutoff $c\ll 1$.

Resetting and repeating this initialization process is necessary to
perform fine tuning and refinement of the weights. A flat histogram
after a simulation in which $w\p{E}$ has changed does not guarantee
that simulating with the final weights will likewise result in a flat
histogram. Eventually, when $f-1<c\ll 1$, i.e. $f\approx 1$, the
initialization process no longer makes any appreciable modifications
to the weights $w\p{E}$, so that upon achieving a flat histogram one
can be confident that a regular simulation which fixes the current
weights will yield similar results.

Interestingly, the flatness condition $C\sp{H}$ for updating $f$ and
restarting the process should not be too stringent: if the histogram
is completely flat, say $H\p{E}=H_0$ for all $E$, then the weights
$w\p{E}$ have all been modified by the same factor of $f^{H_0}$.
Remembering the scale-invariance of $w\p{E}$, this result would mean
that the weights have not actually changed at all! Similarly, too lax
of a flatness condition will prompt the initialization process to
\begin{enumerate*}[label=\roman*)]
\item reset and decrease $f$ too early, i.e. when the weight array
  could still use some heavy handed adjustment, and
\item quit before one has any reason to think that simulations might
  yield a flat histogram.
\end{enumerate*}
Tuning the flatness condition is thus crucial to the success of the
Wang-Landau method.

\begin{algorithm}[!b]
  \caption{Wang-Landau initialization of weights}
  \label{alg:wang_landau}
  \begin{alg}

  \item Construct an initial typical fluid configuration

  \item Set $k=0$, choose some factor $f_0>1$, and initialize
    $w\p{E}=1$ for all $E$.

  \item Simulate for $n$ iterations. After each move, increment the
    histogram $H\p{E}$ of energy observations and update
    $w\p{E}\leftarrow w\p{E}/f_k$, where $E$ is the current energy of
    the system.
    \label{alg:wang_landau_sim}

  \item If the histogram $H\p{E}$ is not sufficiently flat, i.e. if it
    fails a ``flatness'' condition $C\sp{H}$, return to step
    \ref{alg:wang_landau_sim}, otherwise
    \begin{enumerate}
    \item reset $H\p{E}=0$ for all $E$,
    \item increment $k\leftarrow k+1$,
    \item set $f_k=\sqrt[u]{f_{k-1}}$ for some predetermined $u>1$,
      \label{alg:wang_landau_update}
    \item if $f_k\ge c$ for some predetermined $c$, return to step
      \ref{alg:wang_landau_sim}.
    \end{enumerate}

  \end{alg}
\end{algorithm}

Algorithm \ref{alg:wang_landau} spells out the Wang-Landau
initialization process, which has five free parameters: the number of
iterations $n$ to run in step \ref{alg:wang_landau_sim}, the initial
factor $f_0$, the update factor $u$, the flatness condition $C\sp{H}$,
and the cutoff $c$. The first of these parameters, $n$, controls how
often to check the flatness condition $C\sp{H}$ and run updates, if
necessary. The value of $n$ is not too important, so long as it is
reasonable. Too small value of $n$ will waste computation time
determining $C\sp{H}$, while too large of a value will waste time
simulating when $C\sp{H}$ has already been satisfied. In general, $n$
should be proportional to the computation time of $C\sp{H}$, which
scales with energy range of the simulated system. We therefore use
$n=L$, where $L$ is the number of energy levels for the given
square-well fluid (as discussed in Section \ref{sec:simple_flat}).
Papers in the literature typically neglect providing any heuristic or
formula for $n$, reporting a system-independent $n=10^3$ or
$10^5$\cite{wang_landau, wang_landau_mod}.

The next parameter, $f_0$, controls the initial amount by which to
adjust the weight of the current energy after each move. As mentioned
in Algorithm \ref{alg:wang_landau}, one should always have $f_0>1$,
but the appropriate value of $f_0$ will generally depend on the system
at hand. Too large a value of $f_0$ will magnify stochastic error in
initialization, while too small a value of $f_0$ will cause
simulations to take an exceedingly long time to converge.  Lacking any
heuristics for assigning $f_0$, literature typically suggests
$f_0=e$\cite{wang_landau}; after some experimentation with the free
parameters in the Wang-Landau method, we used $f_0=e^{1/8}$ (in
practice, one usually works with $\ln f$, making the commonly
suggested value $\ln f_0=1$ and our own $\ln f_0=1/8$).

The factor $u$ appearing in step \ref{alg:wang_landau_update} of
Algorithm \ref{alg:wang_landau} controls the amount by which to adjust
$f$ when the histogram $H\p{E}$ is sufficiently flat. In principle,
one need not take a root of $f$ at each reset: the general idea behind
Algorithm \ref{alg:wang_landau} requires only $f_k<f_{k-1}$ to
converge (assuming that $C\sp{H}$ can be satisfied) and
$\lim_{k\to\infty}f_k=1$ to terminate. Taking the $u$-th root of $f$
at each increment of $k$ is a natural and convenient iterative means
to satisfy these requirements. Previous papers have used $u=2$, which
we have adopted as well\cite{wang_landau, wang_landau_mod}.

\fixme{Walk through $C\sp{H}$ and $c$. Finish discussion of WL.}

\subsection{The optimized ensemble (OE) method}
\label{sec:optimized_ensemble}

\fixme{Explain, provide algorithm.}

\begin{algorithm}[H]
  \caption{Modifying weights via the optimized ensemble}
  \label{alg:optimized_ensemble}
\end{algorithm}

\subsection{The transition matrix Monte Carlo (TMMC) method}
\label{sec:tmmc}

\fixme{Explain, provide algorithm. This method is sort of our own, but
  based on other work.}

\begin{algorithm}[H]
  \caption{Transition matrix Monte Carlo initialization}
  \label{alg:tmmc}
\end{algorithm}

\subsection{The hybrid OE-TMMC method}
\label{sec:oetmmc}

\fixme{Explain, provide algorithm. This method is sort of our own, but
  based on other work.}

\begin{algorithm}[H]
  \caption{Hybrid OE-TMMC simulation}
  \label{alg:oetmmc}
\end{algorithm}

\subsection{Identifying the energy range of interest}
\label{sec:energy_range}

Histogram methods typically rely on knowledge of the minimum and
maximum energies of the system in question \fixme{how?}, which are
unknown for the square-well fluid. Worse still, some square-well fluid
energies have such an incredibly low density of states that one cannot
reasonably expect to ever observe them via Monte Carlo, biased or
otherwise. As such, we need to identify an energy range of interest,
bounded by minimum and maximum ``important'' energies. \fixme{Rewrite,
  expand}

\subsubsection{State of maximal entropy}
\label{sec:max_entropy}

\fixme{The energy at which the density of states is maximal, i.e. the
  state of maximal entropy, sets an upper bound for the energy range
  in which appropriate weights are not obvious. Explain why.}

\subsubsection{Minimum important energy}
\label{sec:min_energy}

\fixme{Determining the minimum important energy is considerably
  trickier. Punchline: we have to choose a minimum temperature of
  interest, $T_{\t{min}}$, and the minimum important energy is the
  energy at which the slope of the density of states is
  $\exp\p{kT_{\t{min}}/\epsilon}/\epsilon$. This condition for the
  minimum important energy is directly related to canonical weights at
  a temperature of $T_{\t{min}}$.}

\fixme{Mention that in all initializations which simulate for some
  time with unchanging weights (i.e. simple flat), the weight array
  should be set to canonical weights below the minimum important
  energy, as soon as it is identified.}

\fixme{Mention that weights should be set to canonical weights below
  the minimum important energy after all initializations.}

\subsubsection{Computing the density of states}
\label{sec:dos}

\fixme{This section might just become a part of the section on the
  minimum important energy. (\ref{eq:density_of_states}) gives us an
  explicit formula for the density of states if we are confident in
  our energy histogram. We can also compute the density of states from
  the transition matrix, which is sometimes more appropriate to do.}

\subsection{End conditions}
\label{sec:end_conditions}

% \fixme{Talk about how we want method-independent end conditions, if
% possible. Introduce the concept of statistically independent energy
% samples, and explain two measures (one optimistic, and one
% pessimistic) for determining the number of statistically independent
% times we have samples a given energy.}

% \fixme{Talk about two end conditions: a fixed number of times which
% we require to independently sample the minimum important energy, and
% an enforced fractional error in the number of times we have
% independently sampled any energy (at a given minimum temperature).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Discussion}
\label{sec:results}

\subsection{Implementation and overview}
\label{sec:implementation}

\fixme{Talk about the relative difficulty of implementing each method,
  as well as the relative difficulty of optimizing or tuning each
  method. Free parameters are generally bad, and especially so if they
  come without ``obvious'' or prescribed values. Justify our
  modifications to Wang-Landau by showing that we did not make it
  worse. Include mention of how memory requirements for transition
  matrix Monte Carlo scale poorly with system size, as well as energy
  resolution for systems with continuous energies.}

\subsection{Simulation results and errors}
\label{sec:results_and_errors}

\fixme{Talk about the energy histogram, weights, and density of states
  yielded by the methods. These should all be fairly similar. Talk
  about why this is so.}

\begin{figure}[H]
  \centering
  \caption[Energy histograms]{Energy histograms}
  \label{fig:histograms}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Energy weights]{Energy weights}
  \label{fig:weights}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Normalized densities of states]{Normalized densities of
    states}
  \label{fig:density_of_states}
\end{figure}

\fixme{Talk about initialization times, and how they scale with system
  size}

\begin{figure}[H]
  \centering
  \caption[Initialization iterations vs. system size]{Initialization
    iterations vs. system size}
  \label{fig:scaling}
\end{figure}

\fixme{Discuss energy initialization times, sampling rates, and errors
  in computed thermodynamic properties. Explain that errors are given
  relative to a simulation which ran much longer, and talk about how
  we ran that simulation (which method did we use?). Show that all
  methods converge on the same answer as they simulate for longer.}

\begin{figure}[H]
  \centering
  \caption[Optimistic energy sampling rates]{Optimistic energy
    sampling rates}
  \label{fig:opt_sample_rate}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Pessimistic energy sampling rates]{Pessimistic energy
    sampling rates}
  \label{fig:pes_sample_rate}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Initialization iterations vs. data quality]{Initialization
    iterations vs. data quality}
  \label{fig:quality}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Specific internal energy]{Specific internal energy}
  \label{fig:internal_energy}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Specific heat capacity]{Specific heat capacity}
  \label{fig:heat_capacity}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

\fixme{What have we learned? What recommendations can we make? When is
  it appropriate to use which histogram methods?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nocite{*} \bibliography{thesis}

\end{document}