### 2016-04-18 papers/thesis-vischer/abstract.tex

\thispagestyle{plain}
\begin{center}
    \Large
    \textbf{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}

    \vspace{0.8cm}
    \large
    A Computational Project

    \vspace{0.8cm}
    \textbf{Brenden Vischer}

    \vspace{1.2cm}
    \textbf{Abstract}
\end{center}

You take a box, stuff it with spheres, move them around, make it bigger, do it again. Win grant.
\clearpage