### 2016-04-18 papers/thesis-vischer/kinetic_partition.tex

\documentclass[10pt]{article}

%%% standard header
\usepackage[margin=1in]{geometry} % one inch margins
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancyplain} % page formatting style
\usepackage{hyperref} % for linking references
\newcommand{\psl}{6pt} % store spacing as command for consistency
\setlength{\parindent}{0cm} % don't indent new paragraphs...
\parskip \psl % ... place a space between paragraphs instead
\usepackage[inline]{enumitem} % include for \setlist{}, use below
%\setlist{nolistsep} % more compact spacing between environments
\setlist[itemize]{leftmargin=*} % nice margins for itemize ...
\setlist[enumerate]{leftmargin=*} % ... and enumerate environments
\frenchspacing % add a single space after a period
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

\renewcommand{\headrulewidth}{0.5pt} % horizontal line in header
\lhead{Brenden Vischer}
\chead{March 9th, 2015}
\rhead{Kinetic Energy Partition Function}

\begin{document}
\begin{center}
  \large \bf Derivation of Kinetic Contribution to Partition Function
\end{center}

\section*{Partition Function}