### 2016-05-31 papers/thesis-vischer/abstract.tex

\thispagestyle{plain}
\begin{center}
    \Large
    \textbf{Free Energy Decomposition of the Critical Square-Well Liquid Using a Renormalization Group Method}

    \vspace{0.8cm}
    \large
    A Computational Project

    \vspace{0.8cm}
    \textbf{Brenden Vischer}

    \vspace{1.2cm}
    \textbf{Abstract}
\end{center}

\clearpage