### 2017-01-05 papers/thesis-scheirer/final/project report/Project_tex/abstract.tex

\thispagestyle{plain}
\begin{center}
	\Large
	\textbf{Exploring Phase Equilibrium with Statistical non-Associating Fluid Theory:}

	\vspace{0.4cm}
	\large
	A Generalized Renormalization Group Theory Approach.

	\vspace{0.4cm}
	\textbf{Ryan M. Scheirer}

	\vspace{0.9cm}
	\textbf{Abstract}
\end{center}

Through the use of computational modeling we seek an accurate prediction of liquid-vapor coexistence curves.  At the moment, phase equilibrium far from the critical point can be modeled using statistical non-associating fluid theory (SnAFT).  However, our goal is to create better phase coexistence curves near the critical point by incorporating generalized renormalization group theory (GRGT) with SnAFT.  Within GRGT there is an arbitrary parameter "L".  We seek to fit GRGT to SnAFT far from the critical point in such a way that "L" will no longer be fit to the critical behavior.