### 2018-06-29 papers/grand/paper.tex
add embryonic 2d grand paper

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommand{\tssays}[1]{{\color{red} [\cyan{Tanner:} \emph{#1}]}}

\begin{document}
\title{Two-dimensional Broad Histogram Monte Carlo}

\author{Jordan K. Pommerenck}
\author{Cade MIDDLE INITIAL? Trotter}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We present several histogram methods and compare the performance and
  efficiency at treating the square-well fluid.
\end{abstract}

\maketitle

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}