### 2018-07-23 papers/ising/ising-sad.tex
move ising to new paper directory

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for in-line fractions
\usepackage{harpoon}% <---

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommand{\tssays}[1]{{\color{red} [\cyan{Tanner:} \emph{#1}]}}

\newcommand{\Int}{\int}
\newcommand\Emin{E_{\min}}
\newcommand\Nmax{N_{\max}}
\newcommand\Tmin{T_{\min}}
\newcommand\mumax{\mu_{\max}}

\begin{document}
\title{SAD Monte Carlo}

\author{Jordan K. Pommerenck}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
We present several histogram methods and compare the performance and
efficiency at treating the Ising model.
\end{abstract}

\maketitle

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}