### 2019-05-20 papers/histogram/ising.tex
working on writing the ising paper.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Ising Sad paper
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
In this work, we finish writing everything before the abstract~\jpsays{FIX AT END}
\end{abstract}

\maketitle

\section{Introduction}
\jpsays{this needs a total rewrite}
In the early 2000s, Wang and Landau introduced a novel flat-histogram
Monte-Carlo algorithm that accurately determined the density of states
(DOS) for a statistical
system~\cite{wang2001determining,wang2001efficient}. For all of its
power, the method unfortunately requires a priori knowledge of several
user-defined parameters. Thus, for any given system under study, the
user needs to determine the ideal parameters in order to apply the
method. The Wang-Landau algorithm is also known to violate detailed
balance (although only for brief time intervals)~\cite{yan2003fast,
shell2002generalization}. With the violation of detailed balance,
convergence of the algorithm is not guaranteed. The convergence rate
for WL is well known~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. Modification factors that decrease faster than
$1/t$ lead to non-convergence~\cite{belardinelli2007fast}.  These
findings lead to the formation of the $1/t$-WL algorithm.  The ideal
choice of decreasing the modification factor prevents CPU time from
being wasted by continuing to perform calculations after the error in
the density of states is saturated~\cite{belardinelli2008analysis}.
Liang independently considered whether WL could be treated as a special
case of Stochastic Approximation whose convergence could be
mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
considered a form of Stochastic Approximation Monte Carlo (SAMC).
Unlike WL, SAMC can guarantee convergence (if certain conditions are
met). Despite the added benefit of guaranteed convergence, the method
still has a system specific user-defined variable. Such variables often
create difficulty when applying Monte-Carlo methods across arbitrary
systems.  The SAD (Stochastic Approximation with Dynamic $\gamma$)
method~\jpsays{outlined by Pommerenck et.al include David and my paper
on SAD square well in citations!} is special version of the SAMC
algorithm that dynamically chooses the modification factor rather than
relying on system dependent parameters. SAD shares the same convergence
properties with SAMC while replacing un-physical user-defined parameters
with the algorithms dynamic choice.  In this work, we apply the family
of Monte-Carlo algorithms (WL, 1/t-WL, SAMC, SAD) to the 2D Potts andbinning of discrete energies.

%\section{Discrete Spin Systems} In this work, we compare the
%convergence properties of the family of Weight-Based Broad Histogram
%Methods applied to the Q-State Potts model and Ising model. The two
%classical spin systems are ideal testbeds for comparing the convergence
%of Monte-Carlo methods. The Potts and Ising models have 1st and 2nd
%order phase transitions respectively and will both require energy
%binning due to the discrete energy levels of the systems. Convergence
%scaling for the family of Weight-Based Broad Histogram methods can
%easily be explored on the two non-continuous systems.

%\subsection{Potts Model}
%We test the convergence of several Broad Histogram methods
%on the periodic 2D square lattice ($Q=8$) Potts model with nearest
%neighbor interactions~\cite{landau2004new}. The $N\times N$ spin system can
%take on values of $q_i = 1,2,3...Q$ for the Potts spin at site $i$. We
%can write the Hamiltonian as follows:

%\begin{align}
%\mathcal{H} = \sum_{<i,j>} \delta(q_i,q_j)
%\end{align}

%The 2D ten state (Q=10) Potts Model was used as a test system for the Wang
%Landau method during its inception~\cite{wang2001determining,wang2001efficient}.
%Some of the popularity of the model is due to knowledge of exact solutions
%and access to extensive simulation data~\cite{wu1982potts}.
%\jpsays{
%(1) The latent heat of the system which can be obtained by plotting U vs T
%and looking at how much U drops at Tc.
%(2) Looking at the value for the critical temperature Tc.
%Or it is possible that they clarify in a later work~\cite{landau2004new} how
%to extract the correct density of states (Q = 8) in this work.
%\begin{align}
%\sum_E g_N\left( E\right)=Q^N
%\end{align}
%Also the number of ground states (where $E = -2N$) is $Q$. He then re-scales
%the density of states $g_N\left( E\right)$ and writes
%\begin{align}
%\ln \left[ g_N\left(E\right)\right]=\ln \left[ g\left( E\right)\right]
%-\ln \left[ g\left(E=-2N\right)\right]+\ln\left[ Q \right]
%\end{align}
%He calls this the normalized density of states (for his work Q = 10). The
%transition temperature for a 2D lattice $k_B T_C\left(L\right)$ can be directly
%calculated~\cite{landau2004new} and compared with the known value,~\cite{wu1982potts}
%\begin{align}
%k_B T_C = \frac{1}{\ln\left(1+(Q)^{0.5}\right)}
%\end{align}
%Mention: 1st order phase transition nearest neighbor and Hamiltonian
%}

\subsection{Ising Model}
The 2D Ising spin-lattice system is widely used as a testbed when
benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition}. The 2nd order
phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact} make the
system sufficiently interesting for such theoretical comparisons. It is
also of important note that direct comparison of the other methods can
be made with WL as its original implementation was done on this
system~\cite{wang2001determining,wang2001efficient}.
We test the convergence of several flat histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}.
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the infinite volume limit ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over all enumerated spin sites. Beale showed that for finite
lattices the Density of States could directly be calculated from the partition
function~\cite{beale1996exact}
\begin{align}
Z = \sum_E g(E) e^{-{\beta E}}
\end{align}
where $g(E)$ is the multiplicity of the system which is identical to
$D(E)$. We can estimate the average deviation from the exact solution
using the relation~\cite{schneider2017convergence, shakirov2018convergence,
barash2017control}:
\begin{align}
\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
\end{align}
As per the implementation by Wang and
Landau~\cite{wang2001determining}, the total number of available energy
states are $N-1$.
\jpsays{Also the necessary condition that the ground state is
$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
density of states at low temperatures. If $2^N$ states is the condition
used for the ferromagnetic ising model, the rescaled factor is
determined primarily by the maximum density of states impacing the
accuracy of the energy levels at or near the ground state.}

\subsection{Wang-Landau}
The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
landau2014guide} explores energy space by initially setting $\gamma=1$,
and then decreases $\gamma$ in discrete stages.  We track the number of
moves ending at each energy and store the counts in a histogram.  For a
sufficiently flat energy histogram, $\gamma$ is decreased by a specified factor of
$\frac12$.  Flatness is defined by the ratio between the minimum
value of the histogram and its average value.  If this flatness
reaches a specified threshold (typically 0.8), the $\gamma$ value is
decreased and the histogram is reset to zero. The entire process is repeated
until $\gamma$ reaches a desired cutoff.

%The Wang-Landau approach thus has three parameters that need be
%specified: the factor by which to decrease $\gamma$ when flatness is
%acheived, the flatness criterion, and the cutoff that determines when
%the computation is complete.  In addition, an energy range (or in
%general, a \emph{set} of energies) must be supplied, so that the
%flatness criterion can be defined.

Although the algorithm is both efficient and widely used , it
suffers from a few shortcomings.  Frist, an energy range of interest must be
specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
which may require multiple simulations. Second, while Wang-Landau
converges quickly, it does not in general converge to the true density
of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
and does decrease $\gamma$ so quickly that it will never (for any
cutoff value) decrease the error in the density of states beyond a
given nonzero value.

\subsection{$1/t$-Wang-Landau}
1/t WL here.

\subsection{SAMC}
Stochastic Approximation Monte Carlo here.

\subsection{SAD Algorithm}\label{sec:sad}
Stochastic Approximation w/ Dynamic $\gamma$ here.

\section{Results}

\subsection{The 16 $\times$ 16 Ising model}
Something small...
\subsection{128 $\times$ 128 Ising system}
Something big (WL uses 256 by 256 in their paper but only one simulation and
no comparison while more recent comparison type papers use 128 by 128 as max)...

\section{Conclusion}

\section{Acknowledgements}

\bibliography{ising} % Produces the bibliography via BibTeX.

\end{document}