### 2019-11-20 papers/histogram/sad-referee-response-2.md
read through printed paper and made necessary edits. now working on response.

Dear editor,

We appreciate the referees' careful reading of our paper, and feel that their
feedback has led to a significant improvement.  We have made some significant
edits to the paper in response to their comments.  These edits were focused on
providing a broader background with regard to previous work, increasing clarity,
and general proofreading.  We believe that we have addressed all of the
referees' concerns with the exception of the unreasonable request by the first
referee that we present results on additional new systems.

...

Response to first referee
# -----------------------------------------------------------------------------#
In the first round of review I raised the general concern that this
work might not present "substantial improvements of existing
computational methods, provided they are applicable to significant
physics problems," as required for papers to be published in Physical
Review. While the authors have addressed all of my concerns, the
question the editor asks is "Do you find the authors' response
persuasive?" I have thought about this question a lot and tend to say
no, the response is not persuasive. The authors mostly reiterate their
(strong) opinion on the strength of their work (which every author
would do, of course), but did not really convince me that I was very
mistaken in my initial review. For example, I was wondering if the
authors could show results for more than a test system to prove
applicability to "significant physics problems" (as required by PRE).
The authors just state this would be "unreasonable", arguing based on
precedence by referring to six PRE papers that only study a single
system. I could certainly find another six papers, where multiple
systems were studied with a "new Monte Carlo algorithm" that even
showed new physics results. And so on. But the question remains: Do
the authors show evidence for applicability to "significant physics
problems"? Even though the method looks sound and promising and is
well described, I tend to say no. Even though the authors assume this
to be the case, and they are probably right, this evidence is not
shown in the paper.

On the other hand, it is true that the revised version of the paper
contains significant improvements, "providing a broader background
with regard to previous work, increasing clarity, and general
proofreading." I agree. The paper is very well written. It is also
true that the authors addressed all of my previous concerns in a
satisfactory manner. Except for my main concern stated above.

The authors show results for a square-well fluid system. I assume it
would be straightforward to modify the code to simulate a
Lennard-Jones fluid. Since SAD estimates the density of states it
should be possible to show specific heat curves, for example. The
authors could, for example, show that SAD correctly (and much faster
and/or reliably?) finds and samples low-temperature solid-solid
transitions in specific Lennard-Jones clusters (for example, LJ135; J.
Chem. Phys. 124, 204511 (2006)). If that or similar data would be in
the paper there would be no doubt whatsoever that SAD is a
"substantial improvement of existing computational methods".