### 2019-11-20 papers/histogram/sad-referee-response-2.md
I have written a rough draft of the referee response.

Dear editor,

We appreciate the referees' careful reading of our paper, and feel that their
feedback has led to a significant improvement. We have added new data and a
section for the 31 atom Lennard-Jones cluster. We show specific heat curves for
this test system and show the convergence SAD vs the other method quantitatively
(a quantitative comparison of this kind has not been done in the literature to
our knowledge). We also added a note to the Square-well fluid section
highlighting that this system has not been explored using broad histogram
methods. We believe that we have addressed all of the referees' concerns and per
the first referee's own words 'If that or similar data would be in the paper
there would be no doubt whatsoever that SAD is a "substantial improvement of
existing computational methods".'

...

Response to first referee
# -----------------------------------------------------------------------------#

In the previous review, the first referee acknowledged that  'The paper is very
well written. It is also true that the authors addressed all of my previous
concerns in a satisfactory manner.'

The first referee asked that we simulate (either 'faster and/or reliably')
low-temperature solid-solid transitions in specific Lennard-Jones clusters. The
referee gave LJ135 as an example, but stated that if specific heat capacity
curves were in the paper for LJ135 or 'similar' (we did LJ31 for reasons we
state in the paper) then 'there would be no doubt whatsoever that SAD is a
"substantial improvement of existing computational methods".'
We believe that we have completely satisfied the first referee and thank the referee
for pointing our attention toward simulating Lennard-Jones clusters.

Response to second referee
# -----------------------------------------------------------------------------#
All of the second referee's comments were addressed previously.